### 2011-07-25 paper/On Physics-Based 3D Waypoint Generation/paper.tex
addit!
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1092 f4953187-ecf7-0310-90f3-8d0fedb0aae5

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
%\usepackage{subfigure} % ben: is this package allowed?
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
On Physics-Based 3D Waypoint Generation for Autonomous Exploration in Mobile Robotics
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\author{Benjamin Adler and Jianwei Zhang% <-this % stops a space
\thanks{This work was not supported by any organization}% <-this % stops a space
\thanks{H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
        University of Twente, 7500 AE Enschede, The Netherlands
        {\tt\small h.kwakernaak@autsubmit.com}}%
\thanks{P. Misra is with the Department of Electrical Engineering, Wright State University,
        Dayton, OH 45435, USA
        {\tt\small pmisra@cs.wright.edu}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

This paper presents a novel, physics-based approach to generating waypoints, which are subsequently being used to steer a flying platform in three-dimensional space to map an outdoor environment. Because the algorithm relies on simple physical processes, it is both easy to understand and to implement in combination with traditional data structures. Generation of efficient sensor-trajectories for maximized information gain operates directly on unorganized point-clouds, creating a perfect fit for environment mapping with commonly user LIDAR sensors or time-of-flight cameras.

We also present the algorithm by simulating its performance in a virtual outdoor scenario.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

Automatic model building has always been one of the most important parts of autonomous robotics. Exploration and mapping - as a subclass of this problem - has been the topic of many papers in recent years. The SLAM problem was first solved for two-dimensional mapping scenarios [thrun, freiburg etc.] and later-on expanded to three-dimensional environments [kinect-uav etc].

This paper presents a novel approach on increasing the efficiency of such mapping procedures. While the implementations of SLAM have improved greatly in recent years, one aspect of mapping three-dimensional environments was given comparatively little attention: finding the next best view. Typically, SLAM algorithms have been researched and developed on mobile platforms driving on flat surfaces. This setup does not require a high efficiency in exploration, as it does not inherently impose time constraints. The authors of this paper, however, are implementing exploration and mapping algorithms on a flying platform with a maximum flight time of about 15 minutes. Aiming to explore as much of the environment as possible within this limited time motivates an efficient way of finding sensor-poses (and in extension, -trajectories) that enable sensors to deliver as much information as quickly as possible.

\section{State of the Art}

Since the introduction of Thrun's SLAM approach, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments []. Finding the next best view, though, has never been a part of SLAM problem.

In general, frontier heuristic is popular, but how to do this in point cloud? low-res, physics spheres.

Occupancy grid maps are bound to 2d environments or voxel structures.

Multi-Level surface maps, Height maps don't allow for overhang.

e.g. ``Autonomous Exploration for 3D Map Learning'' by Johoo et al.

\section{Our approach}

Prior to the mapping process, the user creates a bounding volume as a representation of the region-of-interest around the vehicle's starting position, thereby defining the environment that is to be mapped. All generated viewpoints will be constrained to this volume, also ensuring that the UAV does not leave the given area. At initialization, a physics engine is set up with a detection volume at the bottom of the defined bounding volume. Whenever an object passes this area in physics computations, a ``calculate

\begin{figure*}[ht]
  \centering
    \includegraphics[width=0.23\textwidth]{images/expl1}
    \includegraphics[width=0.23\textwidth]{images/expl2}
    \includegraphics[width=0.23\textwidth]{images/expl3}
    \includegraphics[width=0.23\textwidth]{images/process5}
    \caption{Mapping setup; bounding volume is blue, detection volume is red, vehicle position is grey, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red. Second figure shows sampling geometry interacting with pointcloud. Geometry hitting the red volume will be converted to a waypoint at the position it last hit a point from the pointcloud. The last figure shows generated pointcloud and leftover sampling geometry.}
    %\label{expl3}
\end{figure*}

As three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, unorganized point clouds are a very common type of sensor data. Finding the next best view in such a point cloud can be very hard, as the data itself does not supply any information about geometric structures such as corners, edges, surfaces and normals. Instead of trying to generate these details from the point cloud whenever it changes, our algorithm does not require such information.

The pointcloud itself is saved into two customized octrees which differ only in the allowed maximum density of points. One octree is used to capture all geometry for later surface reconstruction and thus allows for a very high density of points. The other octree is used for waypoint generation and only stores points if there are no other points whithin a distance of *dMin*, leading to comparatively little memory consumption. This density limitation is implemented by simple neighbor-queries common in most octree implementations.

Any sensor generating spatial occupancy information can be used, as the algorithm's input consists solely of geometry.

Spatially limited by defined exploration space, other input pre-existing sensor information and vehicle position.

Simulated annealing.

kugeln sind gut wgeen einfacher darstellung im speicher und, weil sie rutschen.

\subsection{3D Model / Data structures}

To operate on non-flat surfaces, we need appropriate data structures. Elevation maps don't work for no overhangs and multiple levels.

point cloud, octree

explain how. bullet, btghostobject

Bilder von fallenden Kugeln, abrutschen.

\subsection{Information Gain}

\begin{figure*}[ht]
  \centering
    \includegraphics[width=0.2\textwidth]{images/process1}
    \includegraphics[width=0.2\textwidth]{images/process2}
    \includegraphics[width=0.2\textwidth]{images/process3}
    \includegraphics[width=0.2\textwidth]{images/process4}
    %\includegraphics[width=0.2\textwidth]{images/process5}
    \caption{Multiple iterations of waypoint generation lead to a continuous scanning at the border between known and unknown environment.}
    %\label{expl3}
\end{figure*}

weil die kugeln, die von der kante fallen, (am meisten) zählen, bekomme ich immer wegpunkte mit dem größten information gain?!

\subsection{Computational complexity}

scalable resolution (sample sphere size / octree density)
simulated annealing length

\section{Experiments}
efficiency / performance
graph zeit vs number of points scanned
scalability durch Kugelgröße / Menge

\section{Outlook}

With the algorithm relying on collision detection between many static and dynamic objects, it is currently limited by the speed of collision detection. In general, physics-simulation in general and collision detection especially tends to lend itself well to parallelization using massively-parallel implementations on GPUs

more efficient with opencl -> from iterative to realtime
kugeln spucken in flugrichtung

collision avoidance! lass die kugeln liegen, fliege drüber, dann wird nichts passieren?!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ACKNOWLEDGMENTS}

The authors gratefully acknowledge the contribution of TAMS.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

\bibitem{c1}
J.G.F. Francis, The QR Transformation I, {\it Comput. J.}, vol. 4, 1961, pp 265-271.

\bibitem{c2}
H. Kwakernaak and R. Sivan, {\it Modern Signals and Systems}, Prentice Hall, Englewood Cliffs, NJ; 1991.

\bibitem{c3}
D. Boley and R. Maier, "A Parallel QR Algorithm for the Non-Symmetric Eigenvalue Algorithm", {\it in Third SIAM Conference on Applied Linear Algebra}, Madison, WI, 1988, pp. A20.

\end{thebibliography}

\end{document}

### 2011-07-25 paper/On Physics-Based 3D Waypoint Generation/sample_new.tex
addit!
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1092 f4953187-ecf7-0310-90f3-8d0fedb0aae5

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Preparation of Papers for IEEE Sponsored Conferences \& Symposia
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\author{Huibert Kwakernaak and Pradeep Misra% <-this % stops a space
\thanks{This work was not supported by any organization}% <-this % stops a space
\thanks{H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
        University of Twente, 7500 AE Enschede, The Netherlands
        {\tt\small h.kwakernaak@autsubmit.com}}%
\thanks{P. Misra is with the Department of Electrical Engineering, Wright State University,
        Dayton, OH 45435, USA
        {\tt\small pmisra@cs.wright.edu}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

These instructions provide basic guidelines for preparing camera-ready (CR)
Proceedings-style papers. This document is itself an example of the
desired layout for CR papers (inclusive of this abstract). The document
contains information regarding desktop publishing format, type sizes, and
type faces. Style rules are provided that explain how to handle equations,
units, figures, tables, references, abbreviations, and acronyms. Sections
are also devoted to the preparation of the references and acknowledgments.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

Your goal is to simulate, as closely as possible, the usual appearance of typeset
 papers. This document provides an example of the desired layout and contains
 information regarding desktop publishing format, type sizes, and type faces.

\subsection{Full-Size Camera-Ready (CR) Copy}

If you have desktop publishing facilities, (the use of a computer to aid
 in the assembly of words and illustrations on pages) prepare your CR paper
  in full-size format, on paper 21.6 x 27.9 cm (8.5 x 11 in or 51 x 66 picas).
  It must be output on a printer (e.g., laser printer) having 300 dots/in, or
  better, resolution. Lesser quality printers, such as dot matrix printers,
   are not acceptable, as the manuscript will not reproduce the desired quality.

\subsubsection{Typefaces and Sizes:} There are many different typefaces and a large
variety of fonts (a complete set of characters in the same typeface, style,
 and size). Please use a proportional serif typeface such as Times Roman,
 or Dutch. If these are not available to you, use the closest typeface you
  can. The minimum typesize for the body of the text is 10 point. The minimum
  size for applications like table captions, footnotes, and text subscripts
  is 8 point. As an aid in gauging type size, 1 point is about 0.35 mm (1/72in).
   Examples are as follows:

\subsubsection{Format:} In formatting your original 8.5" x 11" page, set top and
bottom margins to 25 mm (1 in or 6 picas), and left and right margins
to about 18 mm (0.7 in or 4 picas). The column width is 88 mm (3.5 in or 21 picas).
 The space between the two columns is 5 mm(0.2 in or 1 pica). Paragraph
 indentation is about 3.5 mm (0.14 in or 1 pica). Left- and right-justify your
 columns. Cut A4 papers to 28 cm. Use either one or two spaces between sections,
 and between text and tables or figures, to adjust the column length.
  On the last page of your paper, try to adjust the lengths of the
  two-columns so that they are the same. Use automatic hyphenation and
   check spelling. Either digitize or paste your figures.

\begin{table}
\caption{An Example of a Table}
\label{table_example}
\begin{center}
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{UNITS}

Metric units are preferred for use in IEEE publications in light of their
international readership and the inherent convenience of these units in many fields.
In particular, the use of the International System of Units (SI Units) is advocated.
 This system includes a subsystem the MKSA units, which are based on the
 meter, kilogram, second, and ampere. British units may be used as secondary units
 (in parenthesis). An exception is when British units are used as identifiers in trade,
 such as, 3.5 inch disk drive.

\addtolength{\textheight}{-3cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ADDITIONAL REQUIREMENTS}

\subsection{Figures and Tables}

Position figures and tables at the tops and bottoms of columns.
Avoid placing them in the middle of columns. Large figures and tables
may span across both columns. Figure captions should be below the figures;
 table captions should be above the tables. Avoid placing figures and tables
  before their first mention in the text. Use the abbreviation ``Fig. 1'',
  even at the beginning of a sentence.
Figure axis labels are often a source of confusion.
Try to use words rather then symbols. As an example write the quantity ``Inductance",
 or ``Inductance L'', not just.
 Put units in parentheses. Do not label axes only with units.
 In the example, write ``Inductance (mH)'', or ``Inductance L (mH)'', not just ``mH''.
 Do not label axes with the ratio of quantities and units.
 For example, write ``Temperature (K)'', not ``Temperature/K''.

\subsection{Numbering}

Number reference citations consecutively in square brackets \cite{c1}.
 The sentence punctuation follows the brackets \cite{c2}.
 Refer simply to the reference number, as in \cite{c3}.
 Do not use ``ref. \cite{c3}'' or ``reference \cite{c3}''.
Number footnotes separately in superscripts\footnote{This is a footnote}
Place the actual footnote at the bottom of the column in which it is cited.
Do not put footnotes in the reference list.
Use letters for table footnotes (see Table I).

\subsection{Abbreviations and Acronyms}

Define abbreviations and acronyms the first time they are used in the text,
even after they have been defined in the abstract. Abbreviations such as
IEEE, SI, CGS, ac, dc, and rms do not have to be defined. Do not use
abbreviations in the title unless they are unavoidable.

\subsection{Equations}

Number equations consecutively with equation numbers in parentheses flush
 with the right margin, as in (1). To make your equations more compact
 you may use the solidus (/), the exp. function, or appropriate exponents.
  Italicize Roman symbols for quantities and variables, but not Greek symbols.
   Use a long dash rather then hyphen for a minus sign. Use parentheses to avoid
    ambiguities in the denominator.
Punctuate equations with commas or periods when they are part of a sentence:
$$\Gamma_2 a^2 + \Gamma_3 a^3 + \Gamma_4 a^4 + ... = \lambda \Lambda(x),$$
where $\lambda$ is an auxiliary parameter.

Be sure that the symbols in your equation have been defined before the
equation appears or immediately following.
Use ``(1),'' not ``Eq. (1)'' or ``Equation (1),''
except at the beginning of a sentence: ``Equation (1) is ...''.

   \begin{figure}[thpb]
      \centering
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Inductance of oscillation winding on amorphous
       magnetic core versus DC bias magnetic field}
      \label{figurelabel}
   \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CONCLUSIONS AND FUTURE WORKS}

\subsection{Conclusions}

This is a repeat.
Position figures and tables at the tops and bottoms of columns.
Avoid placing them in the middle of columns. Large figures and tables
may span across both columns. Figure captions should be below the figures;
 table captions should be above the tables. Avoid placing figures and tables
  before their first mention in the text. Use the abbreviation ``Fig. 1'',
  even at the beginning of a sentence.
Figure axis labels are often a source of confusion.
Try to use words rather then symbols. As an example write the quantity ``Inductance",
 or ``Inductance L'', not just.
 Put units in parentheses. Do not label axes only with units.
 In the example, write ``Inductance (mH)'', or ``Inductance L (mH)'', not just ``mH''.
 Do not label axes with the ratio of quantities and units.
 For example, write ``Temperature (K)'', not ``Temperature/K''.

\subsection{Future Works}

This is a repeat.
Position figures and tables at the tops and bottoms of columns.
Avoid placing them in the middle of columns. Large figures and tables
may span across both columns. Figure captions should be below the figures;
 table captions should be above the tables. Avoid placing figures and tables
  before their first mention in the text. Use the abbreviation ``Fig. 1'',
  even at the beginning of a sentence.
Figure axis labels are often a source of confusion.
Try to use words rather then symbols. As an example write the quantity ``Inductance",
 or ``Inductance L'', not just.
 Put units in parentheses. Do not label axes only with units.
 In the example, write ``Inductance (mH)'', or ``Inductance L (mH)'', not just ``mH''.
 Do not label axes with the ratio of quantities and units.
 For example, write ``Temperature (K)'', not ``Temperature/K''.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ACKNOWLEDGMENTS}

The authors gratefully acknowledge the contribution of National Research Organization and reviewers' comments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

\bibitem{c1}
J.G.F. Francis, The QR Transformation I, {\it Comput. J.}, vol. 4, 1961, pp 265-271.

\bibitem{c2}
H. Kwakernaak and R. Sivan, {\it Modern Signals and Systems}, Prentice Hall, Englewood Cliffs, NJ; 1991.

\bibitem{c3}
D. Boley and R. Maier, "A Parallel QR Algorithm for the Non-Symmetric Eigenvalue Algorithm", {\it in Third SIAM Conference on Applied Linear Algebra}, Madison, WI, 1988, pp. A20.

\end{thebibliography}

\end{document}

### 2011-08-08 basestation/basestation.cpp
good news, everyone!
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1093 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@ -31,6 +31,7 @@ BaseStation::BaseStation() : QMainWindow()
     connect(mTcpSocket, SIGNAL(disconnected()), SLOT(slotSocketDisconnected()));
 
     mControlWidget = new ControlWidget(this);
+    connect(this, SIGNAL(vehiclePoseChanged(Pose)), mControlWidget, SLOT(slotUpdatePose(Pose)));
     addDockWidget(Qt::RightDockWidgetArea, mControlWidget);
 
     mWirelessDevice = new WirelessDevice("wlan0");
@@ -43,8 +44,9 @@ BaseStation::BaseStation() : QMainWindow()
     mRtkFetcher = new RtkFetcher(mConnectionDialog->getHostNameRtkBase(), 4001, this);
     connect(mRtkFetcher, SIGNAL(rtkData(QByteArray)), SLOT(slotSendRtkDataToRover(QByteArray)));
 
-    mFlightPlanner = new FlightPlannerPhysics(this, &mVehiclePose, mOctree);
+    mFlightPlanner = new FlightPlannerPhysics(this, mOctree);
     mFlightPlanner->slotSetScanVolume(QVector3D(140, 70, 80), QVector3D(240, 120, 150));
+    connect(this, SIGNAL(vehiclePoseChanged(Pose)), mFlightPlanner, SLOT(slotVehiclePoseChanged(Pose)));
     connect(mControlWidget, SIGNAL(setScanVolume(QVector3D,QVector3D)), mFlightPlanner, SLOT(slotSetScanVolume(QVector3D, QVector3D)));
     connect(mControlWidget, SIGNAL(generateWaypoints()), mFlightPlanner, SLOT(slotGenerateWaypoints()));
 
@@ -298,9 +300,10 @@ void BaseStation::processPacket(QByteArray data)
     }
     else if(packetType == "posechanged")
     {
-        stream >> mVehiclePose;
+        Pose vehiclePose;
+        stream >> vehiclePose;
 
-        mControlWidget->slotUpdatePose(mVehiclePose);
+        emit vehiclePoseChanged(vehiclePose);
     }
     else if(packetType == "currentwaypointshash")
     {

### 2011-08-08 basestation/basestation.h

@@ -34,7 +34,7 @@ class BaseStation : public QMainWindow
 
 private:
     QTcpSocket* mTcpSocket;
-    Pose mVehiclePose;
+    //Pose mVehiclePose;
 
     WirelessDevice* mWirelessDevice;
 
@@ -67,6 +67,9 @@ private:
     QProgressDialog* mProgress;
 
     void processPacket(QByteArray data);
+
+signals:
+    void vehiclePoseChanged(Pose);
 
 private slots:
     void slotAskForConnectionHostNames();

### 2011-08-08 basestation/flightplannerbasic.cpp

@@ -1,6 +1,6 @@
 #include "flightplannerbasic.h"
 
-FlightPlannerBasic::FlightPlannerBasic(QWidget* widget, const Pose * const pose, Octree* pointCloud) : FlightPlannerInterface(widget, pose, pointCloud)
+FlightPlannerBasic::FlightPlannerBasic(QWidget* widget, Octree* pointCloud) : FlightPlannerInterface(widget, pointCloud)
 {
     // The octree is initialized on arrival of the first point, with this point at its center.
     // We do this so we can drop spheres only within the octree's XZ plane.

### 2011-08-08 basestation/flightplannerbasic.h

@@ -17,7 +17,7 @@ class FlightPlannerBasic : public FlightPlannerInterface
     Q_OBJECT
 
 public:
-    FlightPlannerBasic(QWidget* widget, const Pose * const pose, Octree* pointCloud);
+    FlightPlannerBasic(QWidget* widget, Octree* pointCloud);
     ~FlightPlannerBasic();
 
     void insertPoint(LidarPoint* const point);

### 2011-08-08 basestation/flightplannerinterface.cpp

@@ -1,9 +1,9 @@
 #include "flightplannerinterface.h"
 
-FlightPlannerInterface::FlightPlannerInterface(QWidget* widget, const Pose * const pose, Octree* pointCloud) : QObject()
+FlightPlannerInterface::FlightPlannerInterface(QWidget* widget, Octree* pointCloud) : QObject()
 {
     mParentWidget = widget;
-    mVehiclePose = pose;
+//    mVehiclePose = pose;
     mWayPointsAhead = new QList<WayPoint>;
     mWayPointsPassed = new QList<WayPoint>;
     qDebug() << "FlightPlannerInterface c'tor.";
@@ -62,12 +62,27 @@ void FlightPlannerInterface::sortToShortestPath(QList<WayPoint> &wayPoints, cons
     qDebug() << "FlightPlannerInterface::sortToShortestPath(): total distance between" << wayPoints.size() << "points after:" << distanceAfter;
 }
 
-const Pose FlightPlannerInterface::getVehiclePose(void) const
+const Pose FlightPlannerInterface::getLastKnownVehiclePose(void) const
 {
-    return *mVehiclePose;
+    if(mVehiclePoses.size())
+        return mVehiclePoses.last();
+    else
+        return Pose();
 }
 
+void FlightPlannerInterface::slotVehiclePoseChanged(const Pose& pose)
+{
+    mVehiclePoses.append(pose);
+    if(mVehiclePoses.size() > 2) mVehiclePoses.takeFirst();
+}
 
+const QVector3D FlightPlannerInterface::getCurrentVehicleVelocity() const
+{
+    if(mVehiclePoses.size() < 2) return QVector3D();
+
+    const quint32 timeDiffMs = mVehiclePoses.last().timestamp - mVehiclePoses.first().timestamp;
+    return (mVehiclePoses.last().position - mVehiclePoses.first().position) * (1000 / timeDiffMs);
+}
 
 void FlightPlannerInterface::slotWayPointDelete(const quint16& index)
 {

### 2011-08-08 basestation/flightplannerinterface.h

@@ -13,21 +13,22 @@ class FlightPlannerInterface : public QObject
     Q_OBJECT
 protected:
     QVector3D mScanVolumeMin, mScanVolumeMax;
-    const Pose *mVehiclePose;
+    QList<Pose> mVehiclePoses;
     QWidget* mParentWidget;
     QList<WayPoint>* mWayPointsAhead, *mWayPointsPassed;
 
     static void sortToShortestPath(QList<WayPoint> &wayPoints, const QVector3D &currentVehiclePosition);
 
 public:
-    FlightPlannerInterface(QWidget* widget, const Pose * const pose, Octree* pointCloud);
+    FlightPlannerInterface(QWidget* widget, Octree* pointCloud);
     virtual ~FlightPlannerInterface();
 
     // Insert points from the laserscanners. Note that the points might also be inserted
     // into Octree* pointCloud, this might be independent from the flightplanner.
     virtual void insertPoint(LidarPoint* const point) = 0;
 
-    const Pose getVehiclePose(void) const;
+    const Pose getLastKnownVehiclePose(void) const;
+    const QVector3D getCurrentVehicleVelocity() const;
 
     const QList<WayPoint> getWayPoints();
 
@@ -44,6 +45,8 @@ public slots:
     virtual void slotSetScanVolume(const QVector3D min, const QVector3D max);
     virtual void slotGenerateWaypoints() = 0;
     virtual void slotVisualize() const = 0;
+
+    void slotVehiclePoseChanged(const Pose& pose);
 
 signals:
     void wayPointInserted(const quint16& index, const WayPoint& wpt);

### 2011-08-08 basestation/flightplannerphysics.cpp

@@ -1,6 +1,6 @@
 #include "flightplannerphysics.h"
 
-FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, const Pose * const pose, Octree* pointCloud) : FlightPlannerInterface(widget, pose, pointCloud)
+FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, Octree* pointCloud) : FlightPlannerInterface(widget, pointCloud)
 {
     // The octree is initialized on arrival of the first point, with this point at its center.
     // We do this so we can drop spheres only within the octree's XZ plane.
@@ -16,6 +16,7 @@ FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, const Pose * const p
 
     connect(mDialog, SIGNAL(gravityChanged(QVector3D)), SLOT(slotGravityChanged(QVector3D)));
     connect(mDialog, SIGNAL(frictionChanged(float,float)), SLOT(slotFrictionChanged(float,float)));
+    connect(mDialog, SIGNAL(restitutionChanged(float,float)), SLOT(slotRestitutionChanged(float,float)));
 
     // Bullet initialisation.
     mBtBroadphase = new btDbvtBroadphase; //new btAxisSweep3(btVector3(-10000,-10000,-10000), btVector3(10000,10000,10000), 1024); // = new btDbvtBroadphase();
@@ -60,7 +61,7 @@ FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, const Pose * const p
 
     mPointCloudGhostObject = new btPairCachingGhostObject();
     mPointCloudGhostObject->setFriction(mDialog->getFrictionGround());
-//    mLidarFloorGhostObject->setRestitution(?);
+    mPointCloudGhostObject->setRestitution(mDialog->getRestitutionGround());
     mPointCloudGhostObject->setWorldTransform(lidarFloorTransform);
     mPointCloudGhostObject->setCollisionShape(mPointCloudShape);
 //    mLidarFloorGhostObject->setCollisionFlags(btCollisionObject::CF_NO_CONTACT_RESPONSE);
@@ -266,6 +267,16 @@ void FlightPlannerPhysics::slotFrictionChanged(const float& frictionPointCloudGr
     }
 }
 
+void FlightPlannerPhysics::slotRestitutionChanged(const float& restitutionPointCloudGround, const float& restitutionSampleGeometry)
+{
[…]

### 2011-08-08 basestation/flightplannerphysics.h

@@ -18,7 +18,7 @@ class FlightPlannerPhysics : public FlightPlannerInterface
 {
     Q_OBJECT
 public:
-    FlightPlannerPhysics(QWidget* widget, const Pose * const pose, Octree* pointCloud);
+    FlightPlannerPhysics(QWidget* widget, Octree* pointCloud);
     ~FlightPlannerPhysics();
 
     void insertPoint(LidarPoint* const point);
@@ -74,6 +74,7 @@ private slots:
     void slotEmitWayPoints();
     void slotGravityChanged(const QVector3D& gravity);
     void slotFrictionChanged(const float& frictionPointCloudGround, const float& frictionSampleGeometry);
+    void slotRestitutionChanged(const float& restitutionPointCloudGround, const float& restitutionSampleGeometry);
 
 public slots:
     void slotSetScanVolume(const QVector3D min, const QVector3D max);

### 2011-08-08 basestation/flightplannerphysicsdialog.cpp

@@ -18,6 +18,9 @@ FlightPlannerPhysicsDialog::FlightPlannerPhysicsDialog(QWidget *parent) :
 
     connect(ui->mSpinBoxFrictionGround, SIGNAL(valueChanged(double)), SLOT(slotFrictionChanged()));
     connect(ui->mSpinBoxFrictionSampleGeometry, SIGNAL(valueChanged(double)), SLOT(slotFrictionChanged()));
+
+    connect(ui->mSpinBoxRestitutionGround, SIGNAL(valueChanged(double)), SLOT(slotRestitutionChanged()));
+    connect(ui->mSpinBoxRestitutionSampleGeometry, SIGNAL(valueChanged(double)), SLOT(slotRestitutionChanged()));
 }
 
 FlightPlannerPhysicsDialog::~FlightPlannerPhysicsDialog()
@@ -41,6 +44,11 @@ void FlightPlannerPhysicsDialog::slotAppendMessage(const QString& message)
     ui->mTextBrowser->append(message/* + "\n"*/);
 }
 
+bool FlightPlannerPhysicsDialog::useRelativeVelocity() const
+{
+    return ui->mChkBoxRelativeVelocity->isChecked();
+}
+
 void FlightPlannerPhysicsDialog::slotGravityChanged()
 {
     emit gravityChanged(
@@ -57,6 +65,11 @@ void FlightPlannerPhysicsDialog::slotFrictionChanged()
     emit frictionChanged(ui->mSpinBoxFrictionGround->value(),ui->mSpinBoxFrictionSampleGeometry->value());
 }
 
+void FlightPlannerPhysicsDialog::slotRestitutionChanged()
+{
+    emit restitutionChanged(ui->mSpinBoxRestitutionGround->value(),ui->mSpinBoxRestitutionSampleGeometry->value());
+}
+
 float FlightPlannerPhysicsDialog::getFrictionGround() const
 {
     return ui->mSpinBoxFrictionGround->value();
@@ -65,6 +78,16 @@ float FlightPlannerPhysicsDialog::getFrictionGround() const
 float FlightPlannerPhysicsDialog::getFrictionSampleGeometry() const
 {
     return ui->mSpinBoxFrictionSampleGeometry->value();
+}
+
+float FlightPlannerPhysicsDialog::getRestitutionGround() const
+{
+    return ui->mSpinBoxRestitutionGround->value();
+}
+
[…]

### 2011-08-08 basestation/flightplannerphysicsdialog.h

@@ -22,10 +22,13 @@ public:
     float getSampleSphereRadius() const;
     float getFrictionGround() const;
     float getFrictionSampleGeometry() const;
+    float getRestitutionGround() const;
+    float getRestitutionSampleGeometry() const;
     bool visualizationActive() const;
     GenerationType getGenerationType() const;
     quint16 getEmitCount() const;
     QVector3D getEmitVelocity() const;
+    bool useRelativeVelocity() const;
 
 private:
     Ui::FlightPlannerPhysicsDialog *ui;
@@ -33,6 +36,7 @@ private:
 private slots:
     void slotGravityChanged();
     void slotFrictionChanged();
+    void slotRestitutionChanged();
 
 public slots:
     void slotSetProgress(const int& value, const int& min, const int& max);
@@ -44,6 +48,7 @@ signals:
     void processPhysics(bool);
     void gravityChanged(QVector3D);
     void frictionChanged(float,float);
+    void restitutionChanged(float,float);
     void submitWayPoints();
 };
 

### 2011-08-08 basestation/glwidget.cpp

@@ -170,7 +170,8 @@ void GlWidget::paintGL()
 //        qDebug() << "centering around 0/0/0" << otc;
 //    }
 
-    mCamLookAt = mFlightPlanner->getVehiclePose().position;
+    const QVector3D vehiclePosition = mFlightPlanner->getLastKnownVehiclePose().position;
+    mCamLookAt = vehiclePosition;
     QVector3D min, max;
     mFlightPlanner->getScanVolume(min, max);
     mCamLookAt = min + (max - min)/2.0;
@@ -196,7 +197,7 @@ void GlWidget::paintGL()
 //    glEnable(GL_BLEND);							// Enable Blending
     glBlendFunc(GL_SRC_ALPHA, GL_ONE_MINUS_SRC_ALPHA);			// Type Of Blending To Use
     glDisable(GL_LIGHTING);
-    OpenGlUtilities::drawSphere(mFlightPlanner->getVehiclePose().position, 1.0, 20.0, QColor(20,255,20,100));
+    OpenGlUtilities::drawSphere(mFlightPlanner->getLastKnownVehiclePose().position, 1.0, 20.0, QColor(20,255,20,100));
 //    glDisable(GL_BLEND);
     glEnable(GL_LIGHTING);
 
@@ -207,6 +208,8 @@ void GlWidget::paintGL()
 
 
     drawAxes(20, 20, 20, 1.0, 1.0, 0.0);
+
+    drawVehicleVelocity();
     // Draw base plate
 //    glDisable(GL_LIGHTING);
 //    glBlendFunc(GL_SRC_ALPHA,GL_ONE);
@@ -218,6 +221,7 @@ void GlWidget::paintGL()
 //    glVertex3f(-1000, 0, -1000);
 //    glEnd();
 //    glEnable(GL_LIGHTING);
+
 
 
 //    mOctree->drawGl();
@@ -234,6 +238,20 @@ void GlWidget::paintGL()
     emit visualizeNow();
 }
 
+void GlWidget::drawVehicleVelocity() const
+{
+    // draw vehicle velocity as vector
+    const QVector3D vehiclePosition = mFlightPlanner->getLastKnownVehiclePose().position;
+    const QVector3D vehicleVelocity = mFlightPlanner->getCurrentVehicleVelocity();
+    glDisable(GL_LIGHTING);
+    glColor3f(1.0, 0.0, 0.0);
+    glLineWidth(1);
+    glBegin(GL_LINES);
[…]

### 2011-08-08 basestation/glwidget.h

@@ -47,6 +47,7 @@ class GlWidget : public QGLWidget
     void wheelEvent(QWheelEvent *event);
     void zoom(double zoomFactor);
     void drawAxes(const GLfloat& x, const GLfloat& y, const GLfloat& z, const GLfloat& red, const GLfloat& green, const GLfloat& blue) const;
+    void drawVehicleVelocity() const;
 
 public:
     GlWidget(BaseStation *baseStation, Octree* octree, FlightPlannerInterface* flightPlanner);

### 2011-08-08 common/plotwidget.h

@@ -5,6 +5,7 @@
 #include <QFile>
 #include <QLayout>
 #include <QString>
+#include <QTextStream>
 #include <QPalette>
 #include <QDateTime>
 #include <QDockWidget>

### 2011-08-08 common/pose.cpp

@@ -173,10 +173,10 @@ Pose Pose::operator+(const Pose &p) const
                 );
 }
 
-/*const Pose Pose::operator+(const Pose &q1, const Pose &q2)
+const Pose Pose::operator-(const Pose &p) const
 {
-    return Pose();
-}*/
+    return Pose(position - p.position, mYaw - p.mYaw, mPitch - p.mPitch, mRoll - p.mRoll);
+}
 
 
 QDebug operator<<(QDebug dbg, const Pose &pose)

### 2011-08-08 common/pose.h

@@ -1,7 +1,7 @@
 #ifndef POSE_H
 #define POSE_H
 
-#include <QVector2D>
+//#include <QVector2D>
 #include <QVector3D>
 #include <QQuaternion>
 #include <QTime>
@@ -42,7 +42,7 @@ public:
 
     Pose operator*(const float &factor);
 //    Pose operator*(const float &factor);
-//    const Pose operator+(const Pose & p);
+    const Pose operator-(const Pose &p) const;
 
     Pose operator+(const Pose &p) const;
 //    friend inline const Pose operator+(const Pose &q1, const Pose &q2);
@@ -65,17 +65,17 @@ public:
     static float getRollRadians(const QQuaternion& orientation, bool reprojectAxis);
     static float getYawRadians(const QQuaternion& orientation, bool reprojectAxis);
 
-    float getPitchRadians() const {return mPitch;};
-    float getRollRadians() const {return mRoll;};
-    float getYawRadians() const {return mYaw;};
+    float getPitchRadians() const {return mPitch;}
+    float getRollRadians() const {return mRoll;}
+    float getYawRadians() const {return mYaw;}
 
-    float getPitchDegrees() const {return RAD2DEG(mPitch);};
-    float getRollDegrees() const {return RAD2DEG(mRoll);};
-    float getYawDegrees() const {return RAD2DEG(mYaw);};
+    float getPitchDegrees() const {return RAD2DEG(mPitch);}
+    float getRollDegrees() const {return RAD2DEG(mRoll);}
+    float getYawDegrees() const {return RAD2DEG(mYaw);}
 
-    void setPitchRadians(const float& pitch) {mPitch = /*normalizeAngleRadians*/(pitch);};
-    void setRollRadians(const float& roll) {mRoll = /*normalizeAngleRadians*/(roll);};
-    void setYawRadians(const float& yaw) {mYaw = /*normalizeAngleRadians*/(yaw);};
+    void setPitchRadians(const float& pitch) {mPitch = /*normalizeAngleRadians*/(pitch);}
+    void setRollRadians(const float& roll) {mRoll = /*normalizeAngleRadians*/(roll);}
+    void setYawRadians(const float& yaw) {mYaw = /*normalizeAngleRadians*/(yaw);}
 
 signals:
 

### 2011-08-09 paper/On Physics-Based 3D Waypoint Generation/paper.tex
in die luft ge&%$st
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1094 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 
-\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
+%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                           % if you need a4paper
-%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
+\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                           % paper
@@
 Since the introduction of Thrun's SLAM approach, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments [].
-Finding the next best view, though, has never been a part of SLAM problem.
+Finding the next best view, though, has never been a part of the SLAM problem and is thus rarely addressed by these works.
 
-In general, frontier heuristic is popular, but how to do this in point cloud? low-res, physics spheres.
+In general, frontier-based approaches are a popular method to computing new waypoints [1,2,3,4], as near-frontier sensor-poses often promise high information gains.
+Unfortunately, information about frontiers between known and unknown space is hard to generate from three-dimensional datastructures such as pointclouds or octrees.
 
-Occupancy grid maps are bound to 2d environments or voxel structures.
+Occupancy grid maps are often used bound to 2d environments or voxel structures.
 
-Multi-Level surface maps, Height maps don't allow for overhang.
+The next-best-view problem is similar, but does not include the constraints of safe navigation. Extension to art-gallery-problem.
+
+Occupancy grid maps - and in extension elevation maps - are often used to store sensor data in generalized form.
+Although the data structure permits easy traversability computation, storing more complex geometry or overhangs remains impossible, limiting their practical use to two-dimensional environments.
+Furthermore, its uniform grid size requires a global compromise between model quality and memory consumption.
+Multi-Level surface maps, introduced by Triebel et al [triebel], eliminate many of these limitations, while still being constrained to a grid.
 
@@
 \section{Our approach}
-
-Prior to the mapping process, the user creates a bounding volume as a representation of the region-of-interest around the vehicle's starting position, thereby defining the environment that is to be mapped.
-All generated viewpoints will be constrained to this volume, also ensuring that the UAV does not leave the given area.
-At initialization, a physics engine is set up with a detection volume at the bottom of the defined bounding volume.
-Whenever an object passes this area in physics computations, a ``calculate
 
@@
 
-The pointcloud itself is saved into two customized octrees which differ only in the allowed maximum density of points.
+While scanning, the pointcloud itself is saved into two customized octrees which differ only in the maximum density of points.
 One octree is used to capture all geometry for later surface reconstruction and thus allows for a very high density of points.
-The other octree is used for waypoint generation and only stores points if there are no other points whithin a distance of *dMin*, leading to comparatively little memory consumption.
+The other octree is used for collision detection and stores points only if there are no other points whithin a distance of *dMin*, leading to comparatively little memory consumption.
 This density limitation is implemented by simple neighbor-queries common in most octree implementations.
+The exact purpose of this second octree will be explained soon.
+
+Prior to the mapping process, the user creates a bounding volume as a representation of the region-of-interest around the vehicle's starting position, thereby defining the environment that is to be mapped.
+All generated viewpoints will be constrained to this volume, ensuring that the UAV does not leave the given area.
+At initialization, a physics engine is set up with a detection volume below the defined bounding volume.
+The contents of the collision-octree are automatically registered as collision objects int he physics world.
+After the octrees are populated with an initial set of points, the waypoint-generation algorithm is executed as follows:
+
+\begin{enumerate}
+     \item Create and initialize sample geometry: spheres of arbitrary size are created along the bounding volume's top plane.
+     \item Start the physics simulation, allowing the sample geometry to follow the defined gravity.
+Whenever a sample sphere collides with a point registered in the collision octree, that event is saved into a special data structure mapping every sphere to the position of its last collision.
+     \item When a sample geometry collides with the detection volume, it is removed from the simulation and the position of its last collision is recorded as a waypoint.
+\end{enumerate}
+
+Although any geometry can be used for waypoint generation, using spheres yields three important advantages:
+\begin{itemize}
+  \item Spheres can represented only by radius and position, so the overall memory requirement is very low (especially when all spheres share a common radius).
+  \item Collision detection between points and spheres is a process of low computational complexity, even redundantizing the narrowphase in collision detection.
+To detect a collision, it is sufficient to check whether the distance between the sphere's center and the point is smaller than the sphere's radius.
+  \item Most importantly, spheres are ideally suited to slip through smaller ``holes'' in the pointcloud, enabling their detection.
+\end{itemize}
 
@@
 
-kugeln sind gut wgeen einfacher darstellung im speicher und, weil sie rutschen.
-
 \subsection{3D Model / Data structures}
@@
 point cloud, octree
+
+The introduction of a second octree for collision detection is, in theory, unnecessary, as all of the pointcloud could be registered as collision objects in the physics world.
+In practice, however, using a second octree with a much lower point density dramatically reduces the number of collision pairs that have to be checked in every step of the physics simulation.
+This reduction translates to an increased efficiency, making the approach feasible for real-time application on current-generation CPUs.
 
@@
 
-\bibitem{c1}
-J.G.F. Francis, The QR Transformation I, {\it Comput. J.}, vol. 4, 1961, pp 265-271.
+\bibitem{gonzalez-banos2002}
+% SLAM does not address the sensor-placement
+% Extension to the ArtGallery-Problem
+% Sensorkeulen von Kandidatenpunkten berechnen, 2D only
+H\'{e}ctor H. Gonz\'{a}lez-Ba\~{n}os and Jean-Claude Latombe, Navigation Strategies for Exploring Indoor Environments, {\it I. J. Robotic Res.} 21(10-11), 829-848 (2002)
+
+\bibitem{haehnel2004}
+% Pioneer und Sick zu Outdoor Modellen, 16seiten, kein Wort über next-best-view
+Dirk H\"ahnel and Wolfram Burgard and Sebastian Thrun, Learning Compact 3d Models of Indoor and Outdoor Environments with a Mobile Robot, {\it Elsevier Science Special Issue Eurobot} '01, 1-16
+
+\bibitem{strand}
+% Next-best-view centered
+% Grid, 2d only
+Marcus Strand and Rüdiger Dillmann, Grid based next best view planning for an autonomous robot in indoor environments
+
+\bibitem{strand2008}
+Marcus Strand and Rüdiger Dillmann, Using an attributed 2D-grid for next-best-view planning on 3D environment data for an autonomous robot
 

### 2011-08-30 paper/On Physics-Based 3D Waypoint Generation/paper.tex
zeugs
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1097 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 While the implementations of SLAM have improved greatly in recent years, one aspect of mapping three-dimensional environments was given comparatively little attention: finding the next best view.
-Typically, SLAM algorithms have been researched and developed on mobile platforms driving on flat surfaces.
+Typically, SLAM algorithms have been researched and developed on mobile platforms driving on flat surfaces[TODO].
 This setup does not require a high efficiency in exploration, as it does not inherently impose time constraints.
@@
 
-Since the introduction of Thrun's SLAM approach, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments [].
+Since the introduction of Thrun's SLAM approach, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments [...].
 Finding the next best view, though, has never been a part of the SLAM problem and is thus rarely addressed by these works.
 
-In general, frontier-based approaches are a popular method to computing new waypoints [1,2,3,4], as near-frontier sensor-poses often promise high information gains.
+Generating safely reachable waypoints from previously acquired sensor data is an extension of next-best-view problem, as it does not include the constraints of safe navigation.
+In general, frontier-based approaches are a popular method to compute new waypoints [1,2,3,4] as sensor-poses from areas between known and unknown environment offer a good compromise between safe reachability and high information gain.
 Unfortunately, information about frontiers between known and unknown space is hard to generate from three-dimensional datastructures such as pointclouds or octrees.
 
-Occupancy grid maps are often used bound to 2d environments or voxel structures.
-
-The next-best-view problem is similar, but does not include the constraints of safe navigation. Extension to art-gallery-problem.
+\subsection{Data structures}
 
 Occupancy grid maps - and in extension elevation maps - are often used to store sensor data in generalized form.
-Although the data structure permits easy traversability computation, storing more complex geometry or overhangs remains impossible, limiting their practical use to two-dimensional environments.
+Although the data structure permits easy traversability computation and frontier detection, storing more complex geometry or overhangs remains impossible, limiting its practical use to two-dimensional environments.
 Furthermore, its uniform grid size requires a global compromise between model quality and memory consumption.
@@
 
-e.g. ``Autonomous Exploration for 3D Map Learning'' by Johoo et al.
+To operate on non-flat surfaces, we need appropriate data structures. Elevation maps don't work for no overhangs and multiple levels.
+
+The introduction of a second octree for collision detection is, in theory, unnecessary, as all of the pointcloud could be registered as collision objects in the physics world.
+In practice, however, using a second octree with a much lower point density dramatically reduces the number of collision pairs that have to be checked in every step of the physics simulation.
+This reduction translates to an increased efficiency, making the approach feasible for real-time application on current-generation CPUs.
+
+explain how. bullet, btghostobject
+
+Bilder von fallenden Kugeln, abrutschen.
 
@@
 As three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, unorganized point clouds are a very common type of sensor data.
-Finding the next best view in such a point cloud can be very hard, as the data itself does not supply any information about geometric structures such as corners, edges, surfaces and normals.
+Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input.
+Finding the next best view in such data be very hard, as the data itself does not supply any information about geometric structures such as corners, edges, surfaces and normals.
 Instead of trying to generate these details from the point cloud whenever it changes, our algorithm does not require such information.
@@
   \item Spheres can represented only by radius and position, so the overall memory requirement is very low (especially when all spheres share a common radius).
-  \item Collision detection between points and spheres is a process of low computational complexity, even redundantizing the narrowphase in collision detection.
+  \item Collision detection between points and spheres is a process of low computational complexity, even redundantizing the midphase in collision detection.
 To detect a collision, it is sufficient to check whether the distance between the sphere's center and the point is smaller than the sphere's radius.
@@
 
-Any sensor generating spatial occupancy information can be used, as the algorithm's input consists solely of geometry.
+%Spatially limited by defined exploration space, other input pre-existing sensor information and vehicle position.
 
-Spatially limited by defined exploration space, other input pre-existing sensor information and vehicle position.
-
-Simulated annealing.
-
-\subsection{3D Model / Data structures}
-
-To operate on non-flat surfaces, we need appropriate data structures. Elevation maps don't work for no overhangs and multiple levels.
-
-point cloud, octree
-
-The introduction of a second octree for collision detection is, in theory, unnecessary, as all of the pointcloud could be registered as collision objects in the physics world.
-In practice, however, using a second octree with a much lower point density dramatically reduces the number of collision pairs that have to be checked in every step of the physics simulation.
-This reduction translates to an increased efficiency, making the approach feasible for real-time application on current-generation CPUs.
-
-explain how. bullet, btghostobject
-
-Bilder von fallenden Kugeln, abrutschen.
+%Simulated annealing.
 
@@
 
+The algorithms complexity derives from the computational effort of the collision detection phase in physics simulation and thus from the number of collision objects.
+Collision detection between $n$ objects in general requires $n(n-1)/2$ collision checks to be performed, leading to a complexity of O($n^2$).
+Optimized algorithms like sort-and-sweep or implementations relying on spatial subdivison may reduce the complexity to O($n$ log $n$) in favorable cases \cite{legrand2007}.
+This initially appears to be a major obstacle to employment of the algorithm outside of simulation, as the pointcloud may grow to several million points during a scan.
+Fortunately, the following optimizations have made the algorithm sufficiently fast for real-time application:
+
+\begin{itemize}
+  \item Given a pointcloud consisting of $n$ points and a set of $m$ sample geometries, the algorithm does not require collision tests between all $n+m$ objects.
+As the $n$ points making up the accumulated scan data are static, the number of required tests is reduced to $\frac{m(m-1)}{2} + n*m$.
+Because $n >> m$, this optimization yields a considerable loss in computational effort.
+  \item When using the sensordata solely for waypoint generation, the density of the pointcloud can be reduced to allow gaps almost the size of the sample geometry's radius.
+That is, if sample spheres are created with a radius of e.g.
+1m, the interacting pointcloud may discard points if there are neighbors within a distance of less than ~2m.
+In practice, this means a reduction of $n$ by two or three dimensions.
+For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neihgbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance, stored in an octree rejecting points that would increase local density above the needed threshold.
+  \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up execution by an order of magnitude compared to calculationi on the CPU.
+\end{itemize}
+
 scalable resolution (sample sphere size / octree density)
@@
 With the algorithm relying on collision detection between many static and dynamic objects, it is currently limited by the speed of collision detection.
-In general, physics-simulation in general and collision detection especially tends to lend itself well to parallelization using massively-parallel implementations on GPUs
+Physics-simulation in general and collision detection especially tends to lend itself well to parallelization using massively-parallel implementations on GPUs
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{ACKNOWLEDGMENTS}
-
-The authors gratefully acknowledge the contribution of TAMS.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.
+%\section{ACKNOWLEDGMENTS}
+%The authors gratefully acknowledge the contribution of TAMS.
 
 \begin{thebibliography}{99}
+
+% ``Autonomous Exploration for 3D Map Learning'' by Johoo et al.
+
+\bibitem{legrand2007}
+% macht broadphase per CUDA, gut erklärt in GPU gems 3, chapter 32
+Scott Le Grand, Broad-Phase Collision Detection with CUDA, {\it GPU Gems 3, } (2007)
 

### 2011-08-30 basestation/basestation.cpp
zeusg
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1098 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@ -460,7 +460,7 @@ void BaseStation::slotSocketError(QAbstractSocket::SocketError socketError)
         QTimer::singleShot(2000, this, SLOT(slotConnectToRover()));
 }
 
-const WayPoint BaseStation::getNextWayPoint(void) const
+/*const WayPoint BaseStation::getNextWayPoint(void) const
 {
     QList<WayPoint> waypoints = mFlightPlanner->getWayPoints();
 
@@ -468,7 +468,7 @@ const WayPoint BaseStation::getNextWayPoint(void) const
         return waypoints.first();
     else
         return WayPoint();
-}
+}*/
 
 /*void BaseStation::slotFlightPlannerProcessing(const QString& text, const quint8& progress)
 {

### 2011-08-30 basestation/basestation.h

@@ -91,7 +91,7 @@ public:
     BaseStation(void);
     ~BaseStation();
 
-    const WayPoint getNextWayPoint(void) const;
+//    const WayPoint getNextWayPoint(void) const;
 };
 
 #endif

### 2011-08-30 basestation/bulletdebugdrawergl.cpp

@@ -1,9 +1,8 @@
 #include "bulletdebugdrawergl.h"
 
-BulletDebugDrawerGl::BulletDebugDrawerGl(/*Ogre::SceneNode *node, btDynamicsWorld *world*/) //:
+BulletDebugDrawerGl::BulletDebugDrawerGl(/*Ogre::SceneNode *node, btDynamicsWorld *world*/) : mDebugOn(true)
 //    mNode(node),
 //    mWorld(world),
-//    mDebugOn(true)
 {
 }
 
@@ -30,7 +29,7 @@ BulletDebugDrawerGl::~BulletDebugDrawerGl()
 
 void BulletDebugDrawerGl::drawLine(const btVector3& from,const btVector3& to,const btVector3& color)
 {
-    qDebug() << "BulletDebugDrawerGl::drawLine";
+//    qDebug() << "BulletDebugDrawerGl::drawLine";
     glLineWidth(1);
     glColor4f(color.x(), color.y(), color.z(), 0.9);
 
@@ -56,7 +55,7 @@ void BulletDebugDrawerGl::drawContactPoint(const btVector3& PointOnB, const btVe
 
 void BulletDebugDrawerGl::drawSphere (btScalar radius, const btTransform &transform, const btVector3 &color)
 {
-    OpenGlUtilities::drawSphere(QVector3D(transform.getOrigin().x(), transform.getOrigin().y(), transform.getOrigin().z()), radius, 5.0, QColor(color.x()*255,color.y()*255,color.z()*255));
+    OpenGlUtilities::drawSphere(QVector3D(transform.getOrigin().x(), transform.getOrigin().y(), transform.getOrigin().z()), radius, 5.0, QColor(color.x()*255,color.y()*255,color.z()*255, 30));
 }
 
 void BulletDebugDrawerGl::drawSphere (const btVector3 &p, btScalar radius, const btVector3 &color)
@@ -123,11 +122,13 @@ void BulletDebugDrawerGl::drawBox(const btVector3 &min, const btVector3 &max, co
 void BulletDebugDrawerGl::reportErrorWarning(const char* warningString)
 {
 //    emit messageWarning(QString(warningString));
+    qDebug() << "BulletDebugDrawerGl::reportErrorWarning():" << warningString;
 }
 
 void BulletDebugDrawerGl::draw3dText(const btVector3& location,const char* textString)
 {
 //    Ogre::LogManager::getSingleton().logMessage(textString);
[…]

### 2011-08-30 basestation/flightplannerphysics.cpp

@@ -32,9 +32,10 @@ FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, Octree* pointCloud) 
       midphase detects which triangles from each trimesh have overlapping AABB and
       narrowphase does collision detection between these triangles. */
     mBtWorld = new btDiscreteDynamicsWorld(mBtDispatcher, mBtBroadphase, mBtSolver, mBtCollisionConfig);
+    mBtWorld->setGravity(btVector3(0,-10,0));
 
-//    mDbgDrawer = new BulletDebugDrawerGl;
-//    mBtWorld->setDebugDrawer(mDbgDrawer);
+    mDbgDrawer = new BulletDebugDrawerGl;
+    mBtWorld->setDebugDrawer(mDbgDrawer);
 
     mTransformLidarPoint.setIdentity();
 
@@ -73,14 +74,16 @@ FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, Octree* pointCloud) 
     mBtWorld->addCollisionObject(mGhostObjectPointCloud, btBroadphaseProxy::SensorTrigger, btBroadphaseProxy::AllFilter & ~btBroadphaseProxy::SensorTrigger);
 
     // Set up vehicle collision avoidance. We create one normal btRigidBody to interact with the world and one ghost object to register collisions
-//    mTransformVehicle.setIdentity();
+    mTransformVehicle.setIdentity();
 //    mShapeVehicle = new btSphereShape(2.0);
-//    mMotionStateVehicle = new btDefaultMotionState;
-    mBodyVehicle = new btRigidBody(btRigidBody::btRigidBodyConstructionInfo(1.0, new btDefaultMotionState, new btSphereShape(2.5), btVector3(0, 0, 0)));
+    mMotionStateVehicle = new btDefaultMotionState;
+    mBodyVehicle = new btRigidBody(btRigidBody::btRigidBodyConstructionInfo(1.0, mMotionStateVehicle, new btSphereShape(2.5), btVector3(0, 0, 0)));
+//    mBtWorld->addRigidBody(mBodyVehicle);
 
     mGhostObjectVehicle = new btPairCachingGhostObject;
     mGhostObjectVehicle->setFriction(10.0);
     mGhostObjectVehicle->setRestitution(0.5);
+    mGhostObjectVehicle->setWorldTransform(mTransformVehicle);
     mGhostObjectVehicle->setCollisionShape(mBodyVehicle->getCollisionShape());
[…]

### 2011-08-30 basestation/flightplannerphysics.h

@@ -47,9 +47,9 @@ private:
     btGhostObject *mGhostObjectDeletionTrigger;
 
     // For vehicle collision avoidance
-//    btTransform mTransformVehicle;
+    btTransform mTransformVehicle;
     btRigidBody *mBodyVehicle;
-//    btDefaultMotionState* mMotionStateVehicle;
+    btDefaultMotionState* mMotionStateVehicle;
 //    btCollisionShape *mShapeVehicle;
     btPairCachingGhostObject *mGhostObjectVehicle;
 
@@ -97,6 +97,9 @@ public slots:
 
     // Overridden from base, create safe path to next waypoint whenever current one was reached.
     void slotWayPointReached(const WayPoint);
+
+    // Overridden from base to move vehicle physics object and check for collisions
+    void slotVehiclePoseChanged(const Pose& pose);
 
     void slotVisualize() const;
 

### 2011-08-30 basestation/glwidget.cpp

@@ -85,7 +85,8 @@ void GlWidget::initializeGL()
 
     glShadeModel(GL_SMOOTH);						// Enable Smooth Shading
 //    glClearColor(0.0f, 0.0f, 0.0f, 0.0f);					// Black Background
-    glClearColor(1.0f, 1.0f, 1.0f, 0.0f);					// White Background
+    glClearColor(0.5f, 0.5f, 0.5f, 0.0f);					// Black Background
+//    glClearColor(1.0f, 1.0f, 1.0f, 0.0f);					// White Background
     glClearDepth(1.0f);							// Depth Buffer Setup
     glHint(GL_LINE_SMOOTH_HINT, GL_NICEST);					// Set Line Antialiasing
 
@@ -202,7 +203,8 @@ void GlWidget::paintGL()
     glEnable(GL_LIGHTING);
 
     // Draw next waypoint
-    OpenGlUtilities::drawSphere(mBaseStation->getNextWayPoint(), 1.0, 20.0, QColor(255,255,255,100));
+    // Don't, let the flightplanners visualize themselves
+//    OpenGlUtilities::drawSphere(mBaseStation->getNextWayPoint(), 1.0, 20.0, QColor(255,255,255,100));
 
 //    glTranslatef(+0.0, +100.0, +0.0);
 

### 2011-09-01 paper/On Physics-Based 3D Waypoint Generation/paper.tex
zeugs
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1099 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 
-\IEEEoverridecommandlockouts                               % This command is only
+%\IEEEoverridecommandlockouts                               % This command is only
                                                            % needed if you want to
@@
 \author{Benjamin Adler and Jianwei Zhang% <-this % stops a space
-\thanks{This work was not supported by any organization}% <-this % stops a space
-\thanks{H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
-        University of Twente, 7500 AE Enschede, The Netherlands
-        {\tt\small h.kwakernaak@autsubmit.com}}%
-\thanks{P. Misra is with the Department of Electrical Engineering, Wright State University,
-        Dayton, OH 45435, USA
-        {\tt\small pmisra@cs.wright.edu}}%
-}
+%\thanks{This work was not supported by any organization}% <-this % stops a space
+%\thanks{H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
+%        University of Twente, 7500 AE Enschede, The Netherlands
+%        {\tt\small h.kwakernaak@autsubmit.com}}%
+%\thanks{P. Misra is with the Department of Electrical Engineering, Wright State University,
+%        Dayton, OH 45435, USA
+%        {\tt\small pmisra@cs.wright.edu}}%
+%}
 
@@
 For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neihgbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance, stored in an octree rejecting points that would increase local density above the needed threshold.
-  \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up execution by an order of magnitude compared to calculationi on the CPU.
+  \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up execution by an order of magnitude compared to calculation on the CPU.
 \end{itemize}

### 2011-09-01 paper/On Physics-Based 3D Waypoint Generation/images/crop.sh

@@ -1,0 +1,1 @@
+find -name "*.png" -exec convert {} -crop 520x340+67+182 cropped/{} \;

### 2011-09-04 paper/On Physics-Based 3D Waypoint Generation/paper.tex
jaja
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1100 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 \section{Our approach}
+
+\subsection{Platform}
+
+To save time during the construction of our platform, we used an "`Okto 2"'-octocopter from the mikrokopter-project \ref{mikrokopter-project} as a base for our vehicle.
+Technically, the central FlightControl (FC) processor-board is connected to eight brushless-motor controllers (BL) via an I2C bus.
+Using on-board gyros and accelerometers, the FC is capable of stabilizing the platform by itself when no other motion-control-commands are coming in from either the connected remote-control-receiver or the serial port.
+The project also offers the NaviControl (NC), an add-on module that sends motion-control-commands (thrust, yaw, pitch, roll as 8-bit signed integers) to steer the platform to waypoints.
+We abandoned the idea of using the NC for our purposes, as its hardware lacks some flexibility and its control-algorithms are not completely open-source.
+
+Instead, we fitted an Intel Atom processor-board to the base.
+This board has a serial connection to the FC, which is used to send motion-control-commands and receive sensor input.
+Furthermore, we mounted a Hokuyo UTM-30LX laserscanner to the platform's front arm such that its scanning-plane is perpendicular to the platform's groundplane.
+
+A Septentrio AsterX2i GPS-receiver has been installed and connected to an XSens MTi IMU, enabling measurement of the platform's pose at 20Hz.
+Due to RTK-GPS, the position is accurate to within 5cm while the orientation shows a maximum error of ~1° for pitch and roll and ~3° for yaw angles.
+To shield the GNSS-receiver from electrostatic interferences caused by the processor board and protect from physical damage, both have been encapsulated into electrically shielded carbon-fiber boxes.
+
+The laserscanner rotates at 2400rpm and provides both falling and rising edges after each single scan.
+This signal is connected the GPS-receiver, which is configured to emit the GNSS-time of the event with microsecond-precision on its USB-connection to the Atom processor-board.
+
+After balancing the platform's extra loads, it exhibits very favorable flight dynamics and drifts very slowly when idling.
+A switch on the remote control can be used to toggle between remote-control and computer-control.
+When in the latter mode, the engines thrust will never exceed the value set on the remote control, providing ideal conditions for testing motion controllers.
+Including payload and battery, the platform weighs TODO grams and requires around 350W of power while hovering, leading to a flight-time between 10 and 15 minutes.
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[scale=1.0]{images/oktokopter}
+  \caption{Rendering of our experimental flying platform}
+  \label{figurelabel}
+\end{figure}
+
+\subsection{Algorithm}
 
@@
 
+\bibitem{mikrokopterproject}
+Mikrokopter project, \url{http://www.mikrokopter.de}, 2011
+
 \bibitem{strand2008}

### 2011-09-05 paper/On Physics-Based 3D Waypoint Generation/paper.tex
test halt
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1102 f4953187-ecf7-0310-90f3-8d0fedb0aae5

--- 2011-09-01 paper/On Physics-Based 3D Waypoint Generation/paper.tex
@@
 %        {\tt\small pmisra@cs.wright.edu}}%
-%}
+}
 
@@
 
-This paper presents a novel, physics-based approach to generating waypoints, which are subsequently being used to steer a flying platform in three-dimensional space to map an outdoor environment.
+This paper presents a novel, physics-based approach to computing waypoints and safe navigation, which are subsequently being used to steer a flying platform in three-dimensional space to map an outdoor environment.
 Because the algorithm relies on simple physical processes, it is both easy to understand and to implement in combination with traditional data structures.
@@
 Exploration and mapping - as a subclass of this problem - has been the topic of many papers in recent years.
-The SLAM problem was first solved for two-dimensional mapping scenarios [thrun, freiburg etc.] and later-on expanded to three-dimensional environments [kinect-uav etc].
+The SLAM problem was first solved for two-dimensional mapping scenarios [nicht-thrun, freiburg etc.] and later-on expanded to three-dimensional environments [kinect-uav etc].
 
@@
 While the implementations of SLAM have improved greatly in recent years, one aspect of mapping three-dimensional environments was given comparatively little attention: finding the next best view.
-Typically, SLAM algorithms have been researched and developed on mobile platforms driving on flat surfaces[TODO].
+Typically, SLAM algorithms have been researched and developed on mobile platforms moving on flat surfaces[TODO].
 This setup does not require a high efficiency in exploration, as it does not inherently impose time constraints.
@@
 
-Since the introduction of Thrun's SLAM approach, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments [...].
+Since the introduction of Thrun's SLAM approach \cite{thrun_invented_slam}, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments [...].
 Finding the next best view, though, has never been a part of the SLAM problem and is thus rarely addressed by these works.
@@
 Occupancy grid maps - and in extension elevation maps - are often used to store sensor data in generalized form.
-Although the data structure permits easy traversability computation and frontier detection, storing more complex geometry or overhangs remains impossible, limiting its practical use to two-dimensional environments.
+Although the data structure permits easy traversability computation and frontier detection, storing more complex geometry or overhangs remains difficult or even impossible, limiting its practical use to two-dimensional environments.
 Furthermore, its uniform grid size requires a global compromise between model quality and memory consumption.
@@
 
-To operate on non-flat surfaces, we need appropriate data structures. Elevation maps don't work for no overhangs and multiple levels.
-
-The introduction of a second octree for collision detection is, in theory, unnecessary, as all of the pointcloud could be registered as collision objects in the physics world.
-In practice, however, using a second octree with a much lower point density dramatically reduces the number of collision pairs that have to be checked in every step of the physics simulation.
-This reduction translates to an increased efficiency, making the approach feasible for real-time application on current-generation CPUs.
-
-explain how. bullet, btghostobject
-
-Bilder von fallenden Kugeln, abrutschen.
+In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based data-structure.
+Octrees are well suited for storing information in non-uniform resolution and the process of storing and retrieving data can easily be executed in parallel, allowing for optimization of parts of our algorithms.
+Besides its position, each point in the octree also stores its brightness (i.e.
+intensity as measured by the laserscanner) and a vector pointing from that point back to the sensor.
+As the platform's measured orientation is less precise that its measured position, this information will be used lateron to assign points recorded from further distance (and thus suffering a greater impact from platform-orientation errors) a lesser weight in the following surface reconstruction.
 
 \section{Our approach}
+
+\subsection{Platform}
+
+\begin{figure}[thpb]
+  \centering
+  \includegraphics[width=0.5\textwidth]{images/oktokopter}
+  \caption{Rendering of our experimental flying platform}
+  \label{oktorendering}
+\end{figure}
+
+To save time during the construction of our platform, we used an ``Okto 2''-octocopter from the mikrokopter-project \cite{mikrokopterproject} as a base for our vehicle.
+Its central FlightControl (FC) processor-board is connected to eight brushless-motor controllers (BL) via an I2C bus.
+Employing its on-board gyros and accelerometers, the FC is programmed to stabilize the platform by itself when no other motion-control-commands are received from either the connected remote-control-receiver or via the serial port.
+The project also developed the NaviControl (NC), an add-on hardware module that sends motion-control-commands (thrust, yaw, pitch and roll each as 8-bit signed integers) to steer the platform to waypoints.
+We abandoned the idea of using the NC for our purposes, as its hardware lacks some flexibility, adds extra weight and its control-algorithms are not open-source.
+
+Instead, we fitted an Intel Atom processor-board to the base.
+This board has a serial connection to the FC, which is used to send motion-control-commands and receive sensor input.
+Furthermore, we mounted a Hokuyo UTM-30LX laserscanner to the platform's front arm such that its scanning-plane is perpendicular to the platform's groundplane (as seen in Fig.
+\ref{oktorendering}).
+
+A Septentrio AsterX2i RTK-GPS receiver has been installed and connected to an XSens MTi IMU, enabling measurement of the platform's pose at 20Hz.
+Due to RTK-GPS, the position is accurate to within 5cm while the orientation shows a maximum error of ~1° for pitch and roll and ~3° for yaw angles.
+To shield the GNSS-receiver from electrostatic interferences caused by the processor board and protect from physical damage, both have been encapsulated into electrically shielded carbon-fiber boxes.
+
+The laserscanner rotates at 2400rpm and provides both falling and rising edges after each single scan.
+This signal is connected to the GPS-receiver, which is configured to emit the GNSS-time of the event with microsecond-precision on its USB-connection to the Atom processor-board.
+With both poses and laserscanner-events being timestamped by the GNSS-receiver, interpolation of the vehicle's pose to the scanner's pose at the time of each scan becomes possible.
+
+% TODO: figure 1: two laserscanners. figure2: one laserscanner
+We initially planned to use one laserscanner with its scan-plane perpendicular to both the platform's groundplane and its forward-vector for scanning the environment and another scanner oriented to have its scan-plane parallel to the vehicle's ground-plane for collision-avoidance (see Fig.
+\ref{lidar-setup-pair}).
+Besides the obvious disadvantages of another 250 grams of payload and ~10W of power consumption, the platform would have had to constantly pitch during flight to ensure reliable obstacle scanning and detection.
+To avoid these drawbacks, we instead mounted a single laserscanner as depicted in Fig.
+\ref{lidar-setup-single}.
+By constantly yawing the platform in flight, our single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the platform does not move further than the laserscanners range within one rotation.
+% TODO: speed < lidar-range * drehgeschwindigkeit ?
+
+After balancing the platform's extra loads, it exhibits very favorable flight dynamics and drifts very slowly when idling.
+A switch on the remote control can be used to toggle between remote-control and computer-control.
+In the latter mode, the engines thrust will never exceed the value set on the remote control, providing ideal conditions for testing motion controllers.
+Including payload and battery, the platform weighs 2250 grams and requires around 350W of power while hovering, leading to a flight-time between 10 and 15 minutes.
+
+\subsection{Algorithm}
 
@@
 The last figure shows generated pointcloud and leftover sampling geometry.}
-    %\label{expl3}
+    \label{explanation_of_process}
 \end{figure*}
@@
 For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neihgbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance, stored in an octree rejecting points that would increase local density above the needed threshold.
+
+  % The introduction of a second octree for collision detection is, in theory, unnecessary, as all of the pointcloud could be registered as collision objects in the physics world.
+In practice, however, using a second octree with a much lower point density dramatically reduces the number of collision pairs that have to be checked in every step of the physics simulation.
+This reduction translates to an increased efficiency, making the approach feasible for real-time application on current-generation CPUs.
+
   \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up execution by an order of magnitude compared to calculation on the CPU.
@@
 
+\bibitem{mikrokopterproject}
+Mikrokopter project, http://www.mikrokopter.de, 2011
+
 \bibitem{strand2008}

### 2011-09-06 paper/On Physics-Based 3D Waypoint Generation/paper.tex
ja
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1104 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 
-This paper presents a novel, physics-based approach to computing waypoints and safe navigation, which are subsequently being used to steer a flying platform in three-dimensional space to map an outdoor environment.
+This paper presents a novel, physics-based approach to computing waypoints and safely navigatable paths between them, which are subsequently being used to steer a flying platform in three-dimensional space to map an outdoor environment.
 Because the algorithm relies on simple physical processes, it is both easy to understand and to implement in combination with traditional data structures.
@@
 Its central FlightControl (FC) processor-board is connected to eight brushless-motor controllers (BL) via an I2C bus.
-Employing its on-board gyros and accelerometers, the FC is programmed to stabilize the platform by itself when no other motion-control-commands are received from either the connected remote-control-receiver or via the serial port.
+Employing its on-board gyros and accelerometers, the FC is programmed to stabilize the platform by itself when no other motion-control-commands are received from either the connected remote-control-receiver or the serial port.
 The project also developed the NaviControl (NC), an add-on hardware module that sends motion-control-commands (thrust, yaw, pitch and roll each as 8-bit signed integers) to steer the platform to waypoints.
-We abandoned the idea of using the NC for our purposes, as its hardware lacks some flexibility, adds extra weight and its control-algorithms are not open-source.
+We abandoned the idea of using the NC for our purposes, as its hardware adds extra weight and its control-algorithms are not open-source.
+Instead, we fitted an Intel Atom processor-board to the platform.
+This board has a serial connection to the FC, which is used to send motion-control-commands and receive sensor input.
 
-Instead, we fitted an Intel Atom processor-board to the base.
-This board has a serial connection to the FC, which is used to send motion-control-commands and receive sensor input.
-Furthermore, we mounted a Hokuyo UTM-30LX laserscanner to the platform's front arm such that its scanning-plane is perpendicular to the platform's groundplane (as seen in Fig.
-\ref{oktorendering}).
-
-A Septentrio AsterX2i RTK-GPS receiver has been installed and connected to an XSens MTi IMU, enabling measurement of the platform's pose at 20Hz.
-Due to RTK-GPS, the position is accurate to within 5cm while the orientation shows a maximum error of ~1° for pitch and roll and ~3° for yaw angles.
+A Septentrio AsterX2i RTK-GNSS (Real Time Kinematic Global Navigation Satellite System) receiver has been installed and connected to an XSens MTi IMU, enabling measurement of the platform's pose at 20Hz.
+Due to RTK-GPS, the position is accurate to within 5cm while the orientation shows a maximum error of $\sim$1° for pitch and roll and $\sim$3° for yaw angles.
 To shield the GNSS-receiver from electrostatic interferences caused by the processor board and protect from physical damage, both have been encapsulated into electrically shielded carbon-fiber boxes.
@@
 This signal is connected to the GPS-receiver, which is configured to emit the GNSS-time of the event with microsecond-precision on its USB-connection to the Atom processor-board.
-With both poses and laserscanner-events being timestamped by the GNSS-receiver, interpolation of the vehicle's pose to the scanner's pose at the time of each scan becomes possible.
+With both poses and laserscanner-events being timestamped by the GNSS-receiver, we can interpolate the vehicle's pose to the scanner's pose at the time of each scan.
 
-% TODO: figure 1: two laserscanners. figure2: one laserscanner
+\begin{figure}[pb]
+  \centering
+  \includegraphics[width=0.48\textwidth]{images/scannerpair}
+  \caption{Setup using two scanners, requiring constant pitching motion}
+  \label{lidar_setup_pair}
+\end{figure}
+
+\begin{figure}[pb]
+  \centering
+  \includegraphics[width=0.48\textwidth]{images/scannersingle}
+  \caption{Setup using single scanner, requiring constant yawing motion}
+  \label{lidar_setup_single}
+\end{figure}
+
 We initially planned to use one laserscanner with its scan-plane perpendicular to both the platform's groundplane and its forward-vector for scanning the environment and another scanner oriented to have its scan-plane parallel to the vehicle's ground-plane for collision-avoidance (see Fig.
-\ref{lidar-setup-pair}).
-Besides the obvious disadvantages of another 250 grams of payload and ~10W of power consumption, the platform would have had to constantly pitch during flight to ensure reliable obstacle scanning and detection.
+\ref{lidar_setup_pair}).
+Besides the obvious disadvantages of another 250 grams of payload and $\sim$10W of power consumption, the platform would have had to constantly pitch during flight to ensure reliable obstacle scanning and detection.
 To avoid these drawbacks, we instead mounted a single laserscanner as depicted in Fig.
-\ref{lidar-setup-single}.
-By constantly yawing the platform in flight, our single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the platform does not move further than the laserscanners range within one rotation.
+\ref{lidar_setup_single}.
+By constantly yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the platform does not move further than the laserscanners range within one rotation.
 % TODO: speed < lidar-range * drehgeschwindigkeit ?
@@
 In the latter mode, the engines thrust will never exceed the value set on the remote control, providing ideal conditions for testing motion controllers.
-Including payload and battery, the platform weighs 2250 grams and requires around 350W of power while hovering, leading to a flight-time between 10 and 15 minutes.
+Including payload and a 5000mAh 4s1p LiPo battery, the platform weighs 2250 grams and requires around 350W of power while hovering, leading to a flight-time of up to 15 minutes.
+
+%TODO \subsection{Motion Control}? Ausrichtung zum Ziel, konstantes drehen?
 
@@
 Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input.
-Finding the next best view in such data be very hard, as the data itself does not supply any information about geometric structures such as corners, edges, surfaces and normals.
-Instead of trying to generate these details from the point cloud whenever it changes, our algorithm does not require such information.
-
-While scanning, the pointcloud itself is saved into two customized octrees which differ only in the maximum density of points.
-One octree is used to capture all geometry for later surface reconstruction and thus allows for a very high density of points.
-The other octree is used for collision detection and stores points only if there are no other points whithin a distance of *dMin*, leading to comparatively little memory consumption.
-This density limitation is implemented by simple neighbor-queries common in most octree implementations.
-The exact purpose of this second octree will be explained soon.
+Finding the next best view in such data can be very hard, as it does not supply any information about geometric structures such as corners, edges, surfaces and normals.
+Instead of trying to generate these details from the point cloud, our algorithm does not require such information.
 
 Prior to the mapping process, the user creates a bounding volume as a representation of the region-of-interest around the vehicle's starting position, thereby defining the environment that is to be mapped.
-All generated viewpoints will be constrained to this volume, ensuring that the UAV does not leave the given area.
+All generated viewpoints will be constrained to this volume, ensuring that the UAV does not leave the confined space.
 At initialization, a physics engine is set up with a detection volume below the defined bounding volume.
-The contents of the collision-octree are automatically registered as collision objects int he physics world.
-After the octrees are populated with an initial set of points, the waypoint-generation algorithm is executed as follows:
+Each point in the pointcloud is automatically registered as a collision object in the physics world.
+After the octree is populated with an initial set of points (i.e.
+right after lift-off), the waypoint-generation algorithm is executed as follows:
 
@@
      \item Start the physics simulation, allowing the sample geometry to follow the defined gravity.
-Whenever a sample sphere collides with a point registered in the collision octree, that event is saved into a special data structure mapping every sphere to the position of its last collision.
-     \item When a sample geometry collides with the detection volume, it is removed from the simulation and the position of its last collision is recorded as a waypoint.
+Whenever a sample geometry instance collides with a point registered in the pointcloud, that event is saved into a special data structure mapping every sample geometry instance to the position of its last collision with the pointcloud.
+     \item When a sample geometry instance collides with the detection volume, it is removed from the simulation and the position of its last collision on the ground-plane is recorded as a waypoint.
 \end{enumerate}
@@
 \begin{itemize}
-  \item Spheres can represented only by radius and position, so the overall memory requirement is very low (especially when all spheres share a common radius).
+  \item Spheres can be represented only by radius and position, reducing the overall memory requirement (especially when all spheres share a common radius).
   \item Collision detection between points and spheres is a process of low computational complexity, even redundantizing the midphase in collision detection.
@@
 
+\subsection{Safe Navigation}
+
+vehicle wird von aktueller position zum wegpunkt gezogen, prall and pointcloud ab, zwischenwegpunkte.
+
 \subsection{Computational complexity}
@@
 That is, if sample spheres are created with a radius of e.g.
-1m, the interacting pointcloud may discard points if there are neighbors within a distance of less than ~2m.
+1m, the interacting pointcloud may discard points if there are neighbors within a distance of less than $\sim$2m.
 In practice, this means a reduction of $n$ by two or three dimensions.
-For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neihgbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance, stored in an octree rejecting points that would increase local density above the needed threshold.
+For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neihgbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance, stored in an octree rejecting points that would increase local density above the required value.
 
@@
 
-  \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up execution by an order of magnitude compared to calculation on the CPU.
+  %While scanning, the pointcloud itself is saved into two customized octrees which differ only in the maximum density of points.
+One octree is used to capture all geometry for later surface reconstruction and thus allows for a very high density of points.
+The other octree is used for collision detection and stores points only if there are no other points whithin a distance of *dMin*, leading to comparatively little memory consumption.
+This density limitation is implemented by simple neighbor-queries common in most octree implementations.
+The exact purpose of this second octree will be explained soon.
+
+  \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up collision detection by an order of magnitude compared to calculation on the CPU.
 \end{itemize}

### 2011-09-07 paper/On Physics-Based 3D Waypoint Generation/paper.tex
jaja
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1105 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 This setup does not require a high efficiency in exploration, as it does not inherently impose time constraints.
-The authors of this paper, however, are implementing exploration and mapping algorithms on a flying platform with a maximum flight time of about 15 minutes.
+The authors of this paper, however, are implementing exploration and mapping algorithms on a flying platform with a maximum flight time of less than 15 minutes.
 Aiming to explore as much of the environment as possible within this limited time motivates an efficient way of finding sensor-poses (and in extension, -trajectories) that enable sensors to deliver as much information as quickly as possible.
@@
 intensity as measured by the laserscanner) and a vector pointing from that point back to the sensor.
-As the platform's measured orientation is less precise that its measured position, this information will be used lateron to assign points recorded from further distance (and thus suffering a greater impact from platform-orientation errors) a lesser weight in the following surface reconstruction.
+As the platform's measured orientation is less precise that its measured position, this vector's length will be used lateron to assign points recorded from further distance (and thus suffering a greater impact from platform-orientation errors) a lesser weight in the following surface reconstruction.
 
@@
 
-To save time during the construction of our platform, we used an ``Okto 2''-octocopter from the mikrokopter-project \cite{mikrokopterproject} as a base for our vehicle.
+To save time during the construction of our platform, we used an ``Okto 2''-octocopter from the mikrokopter-project \cite{mikrokopterproject} as a base for our vehicle (see Fig.
+\ref{oktorendering}).
 Its central FlightControl (FC) processor-board is connected to eight brushless-motor controllers (BL) via an I2C bus.
@@
 The project also developed the NaviControl (NC), an add-on hardware module that sends motion-control-commands (thrust, yaw, pitch and roll each as 8-bit signed integers) to steer the platform to waypoints.
-We abandoned the idea of using the NC for our purposes, as its hardware adds extra weight and its control-algorithms are not open-source.
+We abandoned the idea of using the NC for our purposes, as its hardware adds extra weight and its control-algorithms are neither capable of moving the platform as desired nor open-source.
 Instead, we fitted an Intel Atom processor-board to the platform.
@@
 
+\begin{figure}[b]
+  \centering
+  \includegraphics[width=0.44\textwidth]{images/scannerpair}
+  \caption{Setup using two scanners with collision avoidance requiring permanent pitching motion}
+  \label{lidar_setup_pair}
+\end{figure}
+
 A Septentrio AsterX2i RTK-GNSS (Real Time Kinematic Global Navigation Satellite System) receiver has been installed and connected to an XSens MTi IMU, enabling measurement of the platform's pose at 20Hz.
-Due to RTK-GPS, the position is accurate to within 5cm while the orientation shows a maximum error of $\sim$1° for pitch and roll and $\sim$3° for yaw angles.
+Due to RTK-GPS, the position is accurate to within 5cm while the orientation shows a maximum error of $\sim$1$^\circ$ for pitch and roll and $\sim$3$^\circ$ for yaw angles.
 To shield the GNSS-receiver from electrostatic interferences caused by the processor board and protect from physical damage, both have been encapsulated into electrically shielded carbon-fiber boxes.
 
-The laserscanner rotates at 2400rpm and provides both falling and rising edges after each single scan.
+The laserscanner rotates at 2400rpm and provides both falling and rising edges after each complete scan of its 270$^\circ$ field-of-view.
 This signal is connected to the GPS-receiver, which is configured to emit the GNSS-time of the event with microsecond-precision on its USB-connection to the Atom processor-board.
@@
 
-\begin{figure}[pb]
+\begin{figure}[b]
   \centering
-  \includegraphics[width=0.48\textwidth]{images/scannerpair}
-  \caption{Setup using two scanners, requiring constant pitching motion}
-  \label{lidar_setup_pair}
-\end{figure}
-
-\begin{figure}[pb]
-  \centering
-  \includegraphics[width=0.48\textwidth]{images/scannersingle}
-  \caption{Setup using single scanner, requiring constant yawing motion}
+  \includegraphics[width=0.44\textwidth]{images/scannersingle}
+  \caption{Setup using single scanner, requiring permanent yawing motion}
   \label{lidar_setup_single}
@@
 In the latter mode, the engines thrust will never exceed the value set on the remote control, providing ideal conditions for testing motion controllers.
-Including payload and a 5000mAh 4s1p LiPo battery, the platform weighs 2250 grams and requires around 350W of power while hovering, leading to a flight-time of up to 15 minutes.
+Including payload and a 5000mAh 4s1p LiPo battery, the platform weighs 2250 grams and requires around 350W of power while hovering, leading to a flight-time of up to 12 minutes.
 
@@
   \centering
-    \includegraphics[width=0.23\textwidth]{images/expl1}
-    \includegraphics[width=0.23\textwidth]{images/expl2}
-    \includegraphics[width=0.23\textwidth]{images/expl3}
-    \includegraphics[width=0.23\textwidth]{images/process5}
+    \includegraphics[width=0.45\textwidth]{images/expl1}
+    \includegraphics[width=0.45\textwidth]{images/expl2}
+    \includegraphics[width=0.45\textwidth]{images/expl3}
+    \includegraphics[width=0.45\textwidth]{images/process5}
     \caption{Mapping setup; bounding volume is blue, detection volume is red, vehicle position is grey, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red.
@@
 As three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, unorganized point clouds are a very common type of sensor data.
-Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input.
+Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input from sensors.
 Finding the next best view in such data can be very hard, as it does not supply any information about geometric structures such as corners, edges, surfaces and normals.
-Instead of trying to generate these details from the point cloud, our algorithm does not require such information.
+Fortunately, our algorithm does not require such information.
 
@@
 After the octree is populated with an initial set of points (i.e.
-right after lift-off), the waypoint-generation algorithm is executed as follows:
+right after lift-off), the waypoint-generation algorithm is executed as depicted in Fig.
+\ref{explanation_of_process}:
 
@@
   \centering
-    \includegraphics[width=0.2\textwidth]{images/process1}
-    \includegraphics[width=0.2\textwidth]{images/process2}
-    \includegraphics[width=0.2\textwidth]{images/process3}
-    \includegraphics[width=0.2\textwidth]{images/process4}
+    \includegraphics[width=0.48\textwidth]{images/process1}
+    \includegraphics[width=0.48\textwidth]{images/process2}
+    \includegraphics[width=0.48\textwidth]{images/process3}
+    \includegraphics[width=0.48\textwidth]{images/process4}
     %\includegraphics[width=0.2\textwidth]{images/process5}
-    \caption{Multiple iterations of waypoint generation lead to a continuous scanning at the border between known and unknown environment.}
+    \caption{Multiple iterations of waypoint generation, as seen from the top, lead to continuous scanning at the border between known and unknown environment.}
     %\label{expl3}
@@
 Collision detection between $n$ objects in general requires $n(n-1)/2$ collision checks to be performed, leading to a complexity of O($n^2$).
-Optimized algorithms like sort-and-sweep or implementations relying on spatial subdivison may reduce the complexity to O($n$ log $n$) in favorable cases \cite{legrand2007}.
-This initially appears to be a major obstacle to employment of the algorithm outside of simulation, as the pointcloud may grow to several million points during a scan.
-Fortunately, the following optimizations have made the algorithm sufficiently fast for real-time application:
+Optimized algorithms like sort-and-sweep (\cite{baraff1992}) or implementations relying on spatial subdivison may reduce the complexity to O($n$ log $n$) in favorable cases \cite{legrand2007}.
+This initially appears to be a major obstacle to employment of the algorithm outside of simulation, as the pointcloud may grow to several million points during a scan; after the following optimizations, however, the algorithm has proven to be sufficiently fast for real-time application on current-generation CPUs.
 
@@
 Because $n >> m$, this optimization yields a considerable loss in computational effort.
-  \item When using the sensordata solely for waypoint generation, the density of the pointcloud can be reduced to allow gaps almost the size of the sample geometry's radius.
+
+  \item When using the sensordata solely for waypoint generation, the density of the pointcloud can be reduced to allow gaps almost the size of the sample geometries' radius.
 That is, if sample spheres are created with a radius of e.g.
-1m, the interacting pointcloud may discard points if there are neighbors within a distance of less than $\sim$2m.
-In practice, this means a reduction of $n$ by two or three dimensions.
-For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neihgbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance, stored in an octree rejecting points that would increase local density above the required value.
-
-  % The introduction of a second octree for collision detection is, in theory, unnecessary, as all of the pointcloud could be registered as collision objects in the physics world.
-In practice, however, using a second octree with a much lower point density dramatically reduces the number of collision pairs that have to be checked in every step of the physics simulation.
-This reduction translates to an increased efficiency, making the approach feasible for real-time application on current-generation CPUs.
-
-  %While scanning, the pointcloud itself is saved into two customized octrees which differ only in the maximum density of points.
-One octree is used to capture all geometry for later surface reconstruction and thus allows for a very high density of points.
-The other octree is used for collision detection and stores points only if there are no other points whithin a distance of *dMin*, leading to comparatively little memory consumption.
-This density limitation is implemented by simple neighbor-queries common in most octree implementations.
-The exact purpose of this second octree will be explained soon.
+$r = 1m$, the octree storing that pointcloud may discard points if there are neighbors within a distance of less than $2r$.
+This is implemented by simple neighbor-queries during insertion of candidate points and, while causing a higher computational complexity in that phase, yields a reduction of $n$ by two to three dimensions, dramatically reducing the number of collision pairs that have to be checked in every step of the physics simulation.
+For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neighbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance.
 
@@
 
-scalable resolution (sample sphere size / octree density)
-simulated annealing length
-
-\section{Experiments}
+\section{Results}
 efficiency / performance
@@
 With the algorithm relying on collision detection between many static and dynamic objects, it is currently limited by the speed of collision detection.
-Physics-simulation in general and collision detection especially tends to lend itself well to parallelization using massively-parallel implementations on GPUs
+Physics-simulation in general and collision detection especially lends itself well optimization using massively-parallel implementations on GPUs.
 
-more efficient with opencl -> from iterative to realtime
-kugeln spucken in flugrichtung
-
-collision avoidance! lass die kugeln liegen, fliege drüber, dann wird nichts passieren?!
+We plan to further extend our algorithm by adding another generation-mode for sampling geometry: when created along the bounding volume's top-plane, sample geometry is used to detect unscanned surfaces in all areas of the partially-reconstructed environment.
+A more efficient approach might also emit sampling geometry from the vehicle's current position with a velocity-vector matching the vehicle's course over ground.
+The authors expect that this would generate waypoints closer to the vehicle's position, requiring less distance to be travelled when visiting them.
+Also, as the age of computed waypoints grows, their associated information gain tends to decrease because the surfaces visible from their positions are often already scanned.
+Even worse, waypoints of older age are more susceptible to being enclosed by previously unscanned surface and thus unreachable.
+Although collision avoidance discards these waypoints in a later stage, not enqueueing them in the first place seems to be a better strategy.
 
@@
 Scott Le Grand, Broad-Phase Collision Detection with CUDA, {\it GPU Gems 3, } (2007)
+
+\bibitem{baraff1992}
+David Baraff, Dynamic Simulation of Non-Penetrating Rigid Bodies, Cornell University, pp. 52 (1992)
 

### 2011-09-08 paper/On Physics-Based 3D Waypoint Generation/paper.tex
jaja
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1106 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 
-In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based data-structure.
+In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based \cite{meagher1982} data-structure.
 Octrees are well suited for storing information in non-uniform resolution and the process of storing and retrieving data can easily be executed in parallel, allowing for optimization of parts of our algorithms.
@@
 
-We initially planned to use one laserscanner with its scan-plane perpendicular to both the platform's groundplane and its forward-vector for scanning the environment and another scanner oriented to have its scan-plane parallel to the vehicle's ground-plane for collision-avoidance (see Fig.
+We initially planned to use two laserscanners, one with its scan-plane perpendicular to both the platform's groundplane and its forward-vector for scanning the environment and another scanner oriented to have its scan-plane parallel to the vehicle's ground-plane for collision-avoidance (see Fig.
 \ref{lidar_setup_pair}).
@@
 \ref{lidar_setup_single}.
-By constantly yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the platform does not move further than the laserscanners range within one rotation.
+By constantly yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the world is static and the platform does not move further than the laserscanners range within one rotation.
 % TODO: speed < lidar-range * drehgeschwindigkeit ?
@@
   \centering
-    \includegraphics[width=0.45\textwidth]{images/expl1}
-    \includegraphics[width=0.45\textwidth]{images/expl2}
-    \includegraphics[width=0.45\textwidth]{images/expl3}
-    \includegraphics[width=0.45\textwidth]{images/process5}
+    \includegraphics[width=0.45\textwidth]{images/process_side/1}
+    \includegraphics[width=0.45\textwidth]{images/process_side/2}
+    \includegraphics[width=0.45\textwidth]{images/process_side/3}
+    \includegraphics[width=0.45\textwidth]{images/process_side/4}
     \caption{Mapping setup; bounding volume is blue, detection volume is red, vehicle position is grey, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red.
@@
 
-As three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, unorganized point clouds are a very common type of sensor data.
+Three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, therefore unorganized point clouds are a very common type of sensor data.
 Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input from sensors.
@@
 Prior to the mapping process, the user creates a bounding volume as a representation of the region-of-interest around the vehicle's starting position, thereby defining the environment that is to be mapped.
-All generated viewpoints will be constrained to this volume, ensuring that the UAV does not leave the confined space.
+All generated viewpoints will be constrained to this volume, ensuring that the vehicle does not leave the confined space.
 At initialization, a physics engine is set up with a detection volume below the defined bounding volume.
@@
   \centering
-    \includegraphics[width=0.48\textwidth]{images/process1}
-    \includegraphics[width=0.48\textwidth]{images/process2}
-    \includegraphics[width=0.48\textwidth]{images/process3}
-    \includegraphics[width=0.48\textwidth]{images/process4}
-    %\includegraphics[width=0.2\textwidth]{images/process5}
+    \includegraphics[width=0.45\textwidth]{images/process_top/1}
+    \includegraphics[width=0.45\textwidth]{images/process_top/2}
+    \includegraphics[width=0.45\textwidth]{images/process_top/3}
+    \includegraphics[width=0.45\textwidth]{images/process_top/4}
+    \includegraphics[width=0.45\textwidth]{images/process_top/5}
+    \includegraphics[width=0.45\textwidth]{images/process_top/6}
     \caption{Multiple iterations of waypoint generation, as seen from the top, lead to continuous scanning at the border between known and unknown environment.}
@@
 
+% Octrees
+\bibitem{meagher1982}
+Donald Meagher, Geometric modeling using octree encoding, {\it Computer Graphics and Image Processing}, 19, 2, pp. 129-147, (1982)
+
 \bibitem{legrand2007}

### 2011-09-09 basestation/flightplannerphysics.cpp
jaja
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1107 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@ -8,6 +8,8 @@ FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, Octree* pointCloud) 
 
     mPhysicsProcessingActive = false;
 
+    mFirstSphereHasHitThisIteration = false;
+
     mDialog = new FlightPlannerPhysicsDialog(mParentWidget);
     connect(mDialog, SIGNAL(createSampleGeometry()), SLOT(slotCreateSampleGeometry()));
     connect(mDialog, SIGNAL(deleteSampleGeometry()), SLOT(slotDeleteSampleGeometry()));
@@ -310,7 +312,7 @@ void FlightPlannerPhysics::slotPointInserted(const LidarPoint* lp)
 // NOT in a glBegin()/glEnd() pair.
 void FlightPlannerPhysics::slotVisualize() const
 {
-    mBtWorld->debugDrawWorld();
+//    mBtWorld->debugDrawWorld();
 
     // Draw the scanVolume
     glDisable(GL_LIGHTING);
@@ -321,7 +323,7 @@ void FlightPlannerPhysics::slotVisualize() const
     if(mOctree)
     {
         glPointSize(4);
-        glColor4f(1.0f, 0.0f, 0.0f, 1.0f);
+        glColor4f(1.0f, 0.0f, 0.0f, 0.8f);
 //        glDisable(GL_LIGHTING);
         glBegin(GL_POINTS);
         mOctree->handlePoints(); // => glVertex3f()...
@@ -351,23 +353,23 @@ void FlightPlannerPhysics::slotVisualize() const
     // Draw passed waypoints
     foreach(const WayPoint& wpt, *mWayPointsPassed)
         if(wpt.purpose == WayPoint::SCAN)
-            OpenGlUtilities::drawSphere(wpt, radiusWptScan, 20, QColor(128,255,128, 32));
+            OpenGlUtilities::drawSphere(wpt, radiusWptScan, 20, QColor(128,255,128, 255));
         else
-            OpenGlUtilities::drawSphere(wpt, radiusWptDetour, 20, QColor(128,255,128, 32));
+            OpenGlUtilities::drawSphere(wpt, radiusWptDetour, 20, QColor(128,255,128, 255));
 
-    // Draw fresh waypoints
+    // Draw freshly generated waypoints
     foreach(const WayPoint& wpt, mWayPointsGenerated)
         if(wpt.purpose == WayPoint::SCAN)
[…]

### 2011-09-09 basestation/flightplannerphysics.h

@@ -29,6 +29,8 @@ private:
     BulletDebugDrawerGl* mDbgDrawer;
     bool mPhysicsProcessingActive;
 
+    bool mFirstSphereHasHitThisIteration;
+
     QList<WayPoint> mWayPointsGenerated, mWayPointsDetour;
 
     // testing

### 2011-09-09 basestation/glwidget.cpp

@@ -85,8 +85,8 @@ void GlWidget::initializeGL()
 
     glShadeModel(GL_SMOOTH);						// Enable Smooth Shading
 //    glClearColor(0.0f, 0.0f, 0.0f, 0.0f);					// Black Background
-    glClearColor(0.5f, 0.5f, 0.5f, 0.0f);					// Black Background
-//    glClearColor(1.0f, 1.0f, 1.0f, 0.0f);					// White Background
+//    glClearColor(0.5f, 0.5f, 0.5f, 0.0f);					// Black Background
+    glClearColor(1.0f, 1.0f, 1.0f, 0.0f);					// White Background
     glClearDepth(1.0f);							// Depth Buffer Setup
     glHint(GL_LINE_SMOOTH_HINT, GL_NICEST);					// Set Line Antialiasing
 
@@ -198,7 +198,7 @@ void GlWidget::paintGL()
 //    glEnable(GL_BLEND);							// Enable Blending
     glBlendFunc(GL_SRC_ALPHA, GL_ONE_MINUS_SRC_ALPHA);			// Type Of Blending To Use
     glDisable(GL_LIGHTING);
-    OpenGlUtilities::drawSphere(mFlightPlanner->getLastKnownVehiclePose().position, 1.0, 20.0, QColor(20,255,20,100));
+    OpenGlUtilities::drawSphere(mFlightPlanner->getLastKnownVehiclePose().position, 1.0, 20.0, QColor(0,0,0,200));
 //    glDisable(GL_BLEND);
     glEnable(GL_LIGHTING);
 
@@ -229,7 +229,7 @@ void GlWidget::paintGL()
 //    mOctree->drawGl();
     glDisable(GL_LIGHTING);
     glPointSize(1);
-    glColor4f(.5f, .5f, .5f, 0.9f);
+    glColor4f(.5f, .5f, .5f, 0.7f);
     glBegin(GL_POINTS);
     mOctree->handlePoints();
     glEnd();
@@ -321,7 +321,7 @@ void GlWidget::mouseMoveEvent(QMouseEvent *event)
     rotY = fmod(rotY, 360.0);
     rotZ = fmod(rotZ, 360.0);
 
-//    qDebug() << "rotXYZ:" << rotX << rotY << rotZ;
+    qDebug() << "rotXYZ:" << rotX << rotY << rotZ;
 
     updateGL();
 }
[…]

### 2011-09-09 common/lidarpoint.h

@@ -33,7 +33,6 @@ public:
     // ray might come in handy when we need to find better next-best-
     // shot poses.
     QVector3D position, direction;
-//    Ray ray;
 
     // This tells us the distance between LaserScanner and detected
     // object. Probably helpful because the further an object is away,

### 2011-09-14 paper/On Physics-Based 3D Waypoint Generation/paper.tex
MEHR!
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1108 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 \usepackage{epsfig} % for postscript graphics files
-%\usepackage{subfigure} % ben: is this package allowed?
+\usepackage{subfigure} % ben: is this package allowed?
 %\usepackage{mathptmx} % assumes new font selection scheme installed
@@
 
-%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
-%         \thanks{*Use the $\backslash$thanks command to put information here}\\
-%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
-%         University of Twente\\
-%         7500 AE Enschede, The Netherlands\\
-%         {\tt\small h.kwakernaak@autsubmit.com}}
-%         \hspace*{ 0.5 in}
-%         \parbox{3 in}{ \centering Pradeep Misra**
-%         \thanks{**The footnote marks may be inserted manually}\\
-%        Department of Electrical Engineering \\
-%         Wright State University\\
-%         Dayton, OH 45435, USA\\
-%         {\tt\small pmisra@cs.wright.edu}}
-%}
+\iffalse
+\author{ \parbox{3 in}{\centering Benjamin Adler\\
+         Fachbereich Informatik, Technische Aspekte Multimodaler Systeme (TAMS)\\
+         University of Hamburg\\
+         Hamburg, Germany\\
+         {\tt\small adler@informatik.uni-hamburg.de}}
+         \hspace*{ 0.5 in}
+         \parbox{3 in}{ \centering Jianwei Zhang\\
+         Fachbereich Informatik, Technische Aspekte Multimodaler Systeme (TAMS)\\
+         University of Hamburg\\
+         Hamburg, Germany\\
+         {\tt\small zhang@informatik.uni-hamburg.de}}
+}
+\fi
 
-\author{Benjamin Adler and Jianwei Zhang% <-this % stops a space
-%\thanks{This work was not supported by any organization}% <-this % stops a space
-%\thanks{H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics and Computer Science,
-%        University of Twente, 7500 AE Enschede, The Netherlands
-%        {\tt\small h.kwakernaak@autsubmit.com}}%
-%\thanks{P. Misra is with the Department of Electrical Engineering, Wright State University,
-%        Dayton, OH 45435, USA
-%        {\tt\small pmisra@cs.wright.edu}}%
+\author{Benjamin Adler and Jianwei Zhang\\
+        University of Hamburg\\
+        Fachbereich Informatik,\\
+        Technische Aspekte Multimodaler Systeme (TAMS)\\
+        Hamburg, Germany\\
+        {\tt\small {adler,zhang}@informatik.uni-hamburg.de}%
 }
@@
 
-This paper presents a novel, physics-based approach to computing waypoints and safely navigatable paths between them, which are subsequently being used to steer a flying platform in three-dimensional space to map an outdoor environment.
+This paper presents a novel, physics-based approach to computing waypoints, which are subsequently being used to steer a flying platform in three-dimensional space to map an outdoor environment.
 Because the algorithm relies on simple physical processes, it is both easy to understand and to implement in combination with traditional data structures.
@@
 Exploration and mapping - as a subclass of this problem - has been the topic of many papers in recent years.
-The SLAM problem was first solved for two-dimensional mapping scenarios [nicht-thrun, freiburg etc.] and later-on expanded to three-dimensional environments [kinect-uav etc].
+The SLAM problem was first solved for two-dimensional mapping scenarios and later-on expanded to three-dimensional environments \cite{shen2011}.
 
@@
 While the implementations of SLAM have improved greatly in recent years, one aspect of mapping three-dimensional environments was given comparatively little attention: finding the next best view.
-Typically, SLAM algorithms have been researched and developed on mobile platforms moving on flat surfaces[TODO].
+Typically, SLAM algorithms have been researched and developed on mobile platforms moving on flat surfaces \cite{walthelm2011}, \cite{yamauchi1998}, \cite{joho2007}.
 This setup does not require a high efficiency in exploration, as it does not inherently impose time constraints.
@@
 Aiming to explore as much of the environment as possible within this limited time motivates an efficient way of finding sensor-poses (and in extension, -trajectories) that enable sensors to deliver as much information as quickly as possible.
+While repeatedly passing over the unknown environment in scan-line fashion results in very fast scanning, a better approach needs to be found to ensure scanning of surfaces that are shadowed from those sensor-trajectories.
 
-\section{State of the Art}
+This paper is organized as follows: The next section describes , investigates,
 
-Since the introduction of Thrun's SLAM approach \cite{thrun_invented_slam}, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments [...].
-Finding the next best view, though, has never been a part of the SLAM problem and is thus rarely addressed by these works.
+\section{Related Work}
 
-Generating safely reachable waypoints from previously acquired sensor data is an extension of next-best-view problem, as it does not include the constraints of safe navigation.
-In general, frontier-based approaches are a popular method to compute new waypoints [1,2,3,4] as sensor-poses from areas between known and unknown environment offer a good compromise between safe reachability and high information gain.
-Unfortunately, information about frontiers between known and unknown space is hard to generate from three-dimensional datastructures such as pointclouds or octrees.
+Since the introduction of SLAM, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments \cite{haehnel2004}, TODO.
+Finding the next best view, though, has never been a part of the SLAM problem \cite{gonzalez_banos2002} and is thus rarely addressed by these works.
 
-\subsection{Data structures}
+Generating safely reachable waypoints from previously acquired sensor data is an extension of next-best-view problem (which itself is an extension to the art-gallery problem \cite{orourke1987}), as it does not include the constraints of safe navigation and merely deals with the generation of isolated viewpoints.
+As sensor-poses in areas between known and unknown environment offer a good compromise between safe reachability and high information gain, frontier-based approaches are a popular method to compute new waypoints in two-dimensional space and are well documented in the literature \cite{gonzalez_banos2002}, \cite{strand2009}, \cite{yamauchi1998}, \cite{shade2011}.
+Unfortunately, information about exploration boundaries is hard to generate from three-dimensional pointclouds.
 
-Occupancy grid maps - and in extension elevation maps - are often used to store sensor data in generalized form.
-Although the data structure permits easy traversability computation and frontier detection, storing more complex geometry or overhangs remains difficult or even impossible, limiting its practical use to two-dimensional environments.
-Furthermore, its uniform grid size requires a global compromise between model quality and memory consumption.
-Multi-Level surface maps, introduced by Triebel et al [triebel], eliminate many of these limitations, while still being constrained to a grid.
-
-In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based \cite{meagher1982} data-structure.
-Octrees are well suited for storing information in non-uniform resolution and the process of storing and retrieving data can easily be executed in parallel, allowing for optimization of parts of our algorithms.
-Besides its position, each point in the octree also stores its brightness (i.e.
-intensity as measured by the laserscanner) and a vector pointing from that point back to the sensor.
-As the platform's measured orientation is less precise that its measured position, this vector's length will be used lateron to assign points recorded from further distance (and thus suffering a greater impact from platform-orientation errors) a lesser weight in the following surface reconstruction.
+As our robot is holonomic, the generated waypoints and trajectories can be created with less concern of the platform's ability to catenate them due to motion constraints that other tyes of vehicles (e.g.
+fixed-wing UAVs) previously entailed \cite{hota2009}.
+The common problem of wind, appearing as both opportunity and hindrance, but mostly in the form of further constraints on feasible paths for many UAVs has spawned many contributions in the past \cite{mcneely2007}.
+Even though wind is still a matter of concern for our platform, it does no longer influence the design of path planning or waypoint generation, but only localization and motion control.
 
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{images/oktokopter}
-  \caption{Rendering of our experimental flying platform}
+  \includegraphics[width=0.5\textwidth]{images/oktokopter_photo_small}
+  \caption{The experimental flying platform with mounted GNSS-antenna/receiver, processor-board, laserscanner and IMU.}
   \label{oktorendering}
@@
 Employing its on-board gyros and accelerometers, the FC is programmed to stabilize the platform by itself when no other motion-control-commands are received from either the connected remote-control-receiver or the serial port.
-The project also developed the NaviControl (NC), an add-on hardware module that sends motion-control-commands (thrust, yaw, pitch and roll each as 8-bit signed integers) to steer the platform to waypoints.
+The project also developed the NaviControl (NC), an add-on hardware module that sends motion-control-commands (thrust, yaw, pitch/roll each as 8-bit signed integers) to steer the platform to waypoints.
 We abandoned the idea of using the NC for our purposes, as its hardware adds extra weight and its control-algorithms are neither capable of moving the platform as desired nor open-source.
@@
 A Septentrio AsterX2i RTK-GNSS (Real Time Kinematic Global Navigation Satellite System) receiver has been installed and connected to an XSens MTi IMU, enabling measurement of the platform's pose at 20Hz.
-Due to RTK-GPS, the position is accurate to within 5cm while the orientation shows a maximum error of $\sim$1$^\circ$ for pitch and roll and $\sim$3$^\circ$ for yaw angles.
+Due to RTK-GPS, the position is accurate to within 5cm while the orientation shows a maximum error of $\sim1^\circ$ for pitch and roll and $\sim3^\circ$ for yaw angles.
 To shield the GNSS-receiver from electrostatic interferences caused by the processor board and protect from physical damage, both have been encapsulated into electrically shielded carbon-fiber boxes.
 
-The laserscanner rotates at 2400rpm and provides both falling and rising edges after each complete scan of its 270$^\circ$ field-of-view.
-This signal is connected to the GPS-receiver, which is configured to emit the GNSS-time of the event with microsecond-precision on its USB-connection to the Atom processor-board.
-With both poses and laserscanner-events being timestamped by the GNSS-receiver, we can interpolate the vehicle's pose to the scanner's pose at the time of each scan.
+%Early positioning tests have shown promising results in open spaces. In more crowded areas, where multipath reception becomes a potential problem IR-anti-multipath
+
+The laserscanner rotates at 2400rpm and provides both falling and rising edges after each complete scan of its 270$^\circ$ field of view.
+This signal is connected to the GNSS receiver, which is configured to emit the GPS Time Of Week (TOW) of the event with microsecond-precision on its USB-connection to the Atom processor-board.
+With both poses and laserscanner-events being timestamped by the GNSS-receiver's time system, we can interpolate the vehicle's pose to the scanner's pose at the time of each scan.
 
@@
 Besides the obvious disadvantages of another 250 grams of payload and $\sim$10W of power consumption, the platform would have had to constantly pitch during flight to ensure reliable obstacle scanning and detection.
+Superimposing the pitching motion has proven to result in considerable amounts of work for the motors on the platform's front and back, reducing flight time.
+The non-constant angular pitching speed in this setup also necessitates a more precise interpolation of the vehicle's to the scanner's pose.
+
 To avoid these drawbacks, we instead mounted a single laserscanner as depicted in Fig.
 \ref{lidar_setup_single}.
-By constantly yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the world is static and the platform does not move further than the laserscanners range within one rotation.
+By permanently yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the world is static and the platform does not move further than the laserscanners range within one rotation.
+Given a constant angular yawing-speed, the disadvantage of precise interpolation of the platform's pose becomes less problematic.
 % TODO: speed < lidar-range * drehgeschwindigkeit ?
@@
 
+\subsection{Data structures}
+
+Occupancy grid maps - and in extension elevation maps - are often used to store sensor data in generalized form \cite{shade2011}.
+Although the data structure permits easy traversability computation and frontier detection, storing more complex geometry or overhangs remains difficult or even impossible, limiting its practical use to two-dimensional environments.
+Furthermore, its uniform grid size requires a global compromise between model quality and memory consumption.
+Multi-Level surface maps, introduced by Triebel et al.
+\cite{triebel2006}, eliminate many of these limitations, while still being constrained to a grid.
+
+In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based \cite{meagher1982} data-structure.
+Octrees are well suited for storing information in non-uniform resolution and the process of storing and retrieving data can easily be executed in parallel, allowing for optimization of parts of our algorithms.
+Besides its position, each point in the octree also stores its brightness (i.e.
+intensity as measured by the laserscanner) and a vector pointing from that point back to the sensor.
+As the platform's measured orientation is less precise that its measured position, this vector's length will be used lateron to assign points recorded from further distance (and thus suffering a greater impact from platform-orientation errors) a lesser weight in the following surface reconstruction.
+
+%\begin{figure}[ht]
+%  \centering
+%  \subfigure[Initial setup]{
+%    \label{process_1}
+%    \includegraphics[width=0.4\textwidth]{images/process_side/1}
+%  }
+%  \subfigure[Physics simulation]{
+%    \label{process_2}
+%    \includegraphics[width=0.4\textwidth]{images/process_side/2}
+%  }
+%  \subfigure[Waypoints and residual sampling geometry]{
+%    \label{process_3}
+%    \includegraphics[width=0.4\textwidth]{images/process_side/3}
+%  }
+%  \subfigure[Generated waypoints]{
+%    \label{process_4}
+%    \includegraphics[width=0.4\textwidth]{images/process_side/4}
+%  }
+%  \caption{Waypoint generation, orthogonal projection; the bounding volume is blue, detection volume is red, vehicle position is black, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red.
+Second figure shows sampling geometry interacting with pointcloud.
+Geometry hitting the red volume will be converted to a waypoint at the position it last hit a point from the pointcloud.
+The last figure shows generated waypoints in red for an initial pointcloud population.}
+%  \label{explanation_of_process}
+%\end{figure}
+
 \subsection{Algorithm}
@@
   \centering
-    \includegraphics[width=0.45\textwidth]{images/process_side/1}
-    \includegraphics[width=0.45\textwidth]{images/process_side/2}
-    \includegraphics[width=0.45\textwidth]{images/process_side/3}
-    \includegraphics[width=0.45\textwidth]{images/process_side/4}
-    \caption{Mapping setup; bounding volume is blue, detection volume is red, vehicle position is grey, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red.
+    \includegraphics[width=0.4\textwidth]{images/process_side/1}\hfill
+    \includegraphics[width=0.4\textwidth]{images/process_side/2}\vspace{3mm}
+    \includegraphics[width=0.4\textwidth]{images/process_side/3}\hfill
+    \includegraphics[width=0.4\textwidth]{images/process_side/4}
+    \caption{Waypoint generation, orthogonal projection; the bounding volume is blue, detection volume is red, vehicle position is black, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red.
 Second figure shows sampling geometry interacting with pointcloud.
 Geometry hitting the red volume will be converted to a waypoint at the position it last hit a point from the pointcloud.
-The last figure shows generated pointcloud and leftover sampling geometry.}
+The last figure shows generated waypoints in red for an initial pointcloud population.}
     \label{explanation_of_process}
@@
 
-Three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, therefore unorganized point clouds are a very common type of sensor data.
-Any sensor generating spatial occupancy information can be used, as the pointcloud is the algorithm's only input from sensors.
+Three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, so unorganized point clouds are a very common type of sensor data.
+Any sensor generating spatial occupancy information can be used with our algorithm, as the pointcloud is its only sensory input.
 Finding the next best view in such data can be very hard, as it does not supply any information about geometric structures such as corners, edges, surfaces and normals.
-Fortunately, our algorithm does not require such information.
+In contrast to many other algorithms (\cite{joho2007}, \cite{walthelm2011}), ours does not require such information.
 
 Prior to the mapping process, the user creates a bounding volume as a representation of the region-of-interest around the vehicle's starting position, thereby defining the environment that is to be mapped.
-All generated viewpoints will be constrained to this volume, ensuring that the vehicle does not leave the confined space.
-At initialization, a physics engine is set up with a detection volume below the defined bounding volume.
-Each point in the pointcloud is automatically registered as a collision object in the physics world.
+All generated viewpoints will be constrained to this volume and its vicinity, ensuring that the vehicle does not leave the scene.
+During initialization, a physics engine is set up with a detection volume below the defined bounding volume.
+Each new point in the pointcloud is automatically registered as a collision object in the physics world.
 After the octree is populated with an initial set of points (i.e.
@@
 \begin{enumerate}
-     \item Create and initialize sample geometry: spheres of arbitrary size are created along the bounding volume's top plane.
-     \item Start the physics simulation, allowing the sample geometry to follow the defined gravity.
-Whenever a sample geometry instance collides with a point registered in the pointcloud, that event is saved into a special data structure mapping every sample geometry instance to the position of its last collision with the pointcloud.
-     \item When a sample geometry instance collides with the detection volume, it is removed from the simulation and the position of its last collision on the ground-plane is recorded as a waypoint.
+     \item Create and initialize sample geometry: spheres of user-defined size are spwaned evenly along the bounding volume's top plane.
+Increasing the sample geometry's size will create fewer instances, less computational burden (see section \ref{complexity}) and less waypoints, while decreasing the size will find smaller holes the the sampled surface.
+     \item Start the physics simulation, allowing the sampling geometry to follow the defined gravity.
+Whenever a sample geometry instance collides with a point registered in the pointcloud, that event is saved into a special data structure mapping every sample geometry instance to its position in that moment.
+As this data structure is implemented as a map, later events caused by the same sampling geometry instance replace its previously saved position.
+     \item When a sample geometry instance collides with the detection volume, it is removed from the simulation and the position of its last collision on the ground-plane is recorded as a waypoint-marker.
+     \item Waypoint-markers are converted to waypoints if there are no other waypoints in close vicinity.
+When converting markers to waypoints, their height needs to be adjusted, as waypoint-markers tend to spawn on the scenery's ground.
+The distance $d$, by which markers are lifted before becoming waypoints, is calculated by $d = Range_{lidar} * \frac{SizeSampleGeom_{current}}{SizeSampleGeom_{max}}$.
+By making the distance from laserscanner to the scanned surface depend on the size of the surface's holes, a natural behavior TODO emerges: The vehicle will first scan larger voids in the surface from further away, then, as the sampling geometry's size is reduced in later iterations of the waypoint generation, come closer to fill smaller defects that couldn't be seen from former sensor poses.
 \end{enumerate}
+
+Physics simulation stops when all sampling geometry has either been deleted after hitting the delection volume, is currently in sleep state or has stopped moving entirely.
+Depending on the solver employed in the physics calculations, the weight of the involved bodies and the number of contact pairs between them, the latter criterion cannot always be fulfilled, as numerical instabilities can cause simulated objects to continue moving indefinetely.
+For this reason, we implemented a threshold-based approach, putting slowly-moving bodies in sleep states until they are reactivated by impulses from other bodies.
+This way, termination of the physics computation is guaranteed.
 
@@
 
-\subsection{Information Gain}
-
 \begin{figure*}[ht]
   \centering
-    \includegraphics[width=0.45\textwidth]{images/process_top/1}
-    \includegraphics[width=0.45\textwidth]{images/process_top/2}
-    \includegraphics[width=0.45\textwidth]{images/process_top/3}
-    \includegraphics[width=0.45\textwidth]{images/process_top/4}
-    \includegraphics[width=0.45\textwidth]{images/process_top/5}
-    \includegraphics[width=0.45\textwidth]{images/process_top/6}
-    \caption{Multiple iterations of waypoint generation, as seen from the top, lead to continuous scanning at the border between known and unknown environment.}
-    %\label{expl3}
+    \includegraphics[width=0.3\textwidth]{images/process_top/1}\hfill
+    \includegraphics[width=0.3\textwidth]{images/process_top/2}\hfill
+    \includegraphics[width=0.3\textwidth]{images/process_top/3}\vspace{5mm}
+    \includegraphics[width=0.3\textwidth]{images/process_top/4}\hfill
+    \includegraphics[width=0.3\textwidth]{images/process_top/5}\hfill
+    \includegraphics[width=0.3\textwidth]{images/process_top/6}
+    \caption{As seen from the top, multiple iterations of waypoint generation lead to continuous scanning at the border between known and unknown environment.
+Red waypoints are enqueued for scanning, green waypoints have been passed.
+In the second-to-last figure, sampling geometry of reduced size has allowed finding smaller holes in previously scanned surfaces.}
+    \label{figure_process}
 \end{figure*}
 
-weil die kugeln, die von der kante fallen, (am meisten) zählen, bekomme ich immer wegpunkte mit dem größten information gain?!
-
-\subsection{Safe Navigation}
-
-vehicle wird von aktueller position zum wegpunkt gezogen, prall and pointcloud ab, zwischenwegpunkte.
+%\subsection{Information Gain}
+%To increase the sampled area of 
+%weil die kugeln, die von der kante fallen, (am meisten) zählen, bekomme ich immer wegpunkte mit dem größten information gain?!
 
 \subsection{Computational complexity}
+\label{complexity}
 
@@
 As the $n$ points making up the accumulated scan data are static, the number of required tests is reduced to $\frac{m(m-1)}{2} + n*m$.
-Because $n >> m$, this optimization yields a considerable loss in computational effort.
+Because $n \gg m$, this optimization yields a considerable loss in computational effort.
 
@@
 $r = 1m$, the octree storing that pointcloud may discard points if there are neighbors within a distance of less than $2r$.
-This is implemented by simple neighbor-queries during insertion of candidate points and, while causing a higher computational complexity in that phase, yields a reduction of $n$ by two to three dimensions, dramatically reducing the number of collision pairs that have to be checked in every step of the physics simulation.
+This is implemented by simple neighbor-queries during insertion of candidate points and - while causing a higher computational complexity in that phase - yields a reduction of $n$ by about two dimensions, dramatically reducing the number of collision pairs that have to be checked in every step of the physics simulation.
 For this reason, we use two pointclouds in our work; one is used for surface reconstruction and is stored in an octree which allows for close neighbors (and thus high density) and another pointcloud used for waypoint generation and collision avoidance.
 
-  \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up collision detection by an order of magnitude compared to calculation on the CPU.
+  \item As described in Le Grand's work \cite{legrand2007}, execution of the simulation's broadphase on GPU can speed up collision detection by more than an order of magnitude compared to calculation on the CPU.
 \end{itemize}
@@
 \section{Results}
-efficiency / performance
-graph zeit vs number of points scanned
-scalability durch Kugelgröße / Menge
+
+\begin{figure}[bt]
+  \centering
+    \includegraphics[width=0.48\textwidth]{images/graph_points_vs_time}
+    \caption{Pointcloud size vs.
+flight-time using several waypoint generators.
+Vertical lines are drawn on every round of waypoint generation using our algorithm, executed after all previously generated waypoints have been passed.}
+    \label{graph_points_vs_time}
+\end{figure}
+
+To assess our algorithms performance, we established the number of points stored in the pointcloud for a given flight-time as the primary metric.
+As explained in chapter \ref{complexity}, our octree-implementation allows setting a maximum point-density during construction.
+For the pointcloud storing the surface-reconstruction data, we defined the minimum distance between neighboring points to be $0.1m$.
+Because the number of points stored correlates directly to the scanned surface-area, measuring the number of points stored over time equals measuring the scanned surface area over time in meaning.
+All tests were executed in a simulator where the platform's angular velocity remained constant at $120 ^\circ/s$, while its linear velocity was limited to $2.8 m/s$ during all trials.
+After the scandata was streamed to the basestation using a TCP-connection, it was processed and statistics about flight-time and pointcloud-sizes were logged.
+
+The graph in Fig.
+\ref{graph_points_vs_time} shows results of using a simple, manually planned and collision-free scanline based exploration with trajectories similar to those for optimal complete terrain coverage in \cite{xu2011}.
+The results of scanning along these trajectories can be seen in Fig.
+\ref{cloud_top_scanlines_4} and \ref{cloud_top_scanlines_6} for four and six equidistant passes, respectively.
+As expected, flying more scanlines in the same area results in more points scanned, albeit with a visibly slower rate.
+
+Additionally, the graph's red line depicts our algorithms' behavior: because the pointcloud needs an initial population for waypoint generation, the scanning rate stagnates shortly after launch during the first phase of waypoint generation, then reaches comparable speeds moments later.
+As
+
+\begin{figure}[htbp]
+  \centering
+  \subfigure[4 passes]{
+    \label{cloud_top_scanlines_4}
+    \includegraphics[width=0.22\textwidth]{images/cloud_top_scanlines_4} 
+  }
+  \subfigure[6 passes]{
+    \label{cloud_top_scanlines_6}
+    \includegraphics[width=0.22\textwidth]{images/cloud_top_scanlines_6} 
+  }
+  \caption{Pointclouds resulting from straight scanline passes. Note the unscanned surface in the bottom center shadowed by a roof.}
+  \label{cloud_top_scanlines}
+\end{figure}
 
@@
 
-With the algorithm relying on collision detection between many static and dynamic objects, it is currently limited by the speed of collision detection.
-Physics-simulation in general and collision detection especially lends itself well optimization using massively-parallel implementations on GPUs.
+With the algorithm relying on collision detection between large numbers of static and dynamic objects, it is currently limited by the speed of collision detection.
+Physics-simulation in general and collision detection especially lends itself well to optimization using massively-parallel implementations on GPUs.
+In the next steps, tests and extensions of the algorithm on GPUs are planned.
+Here, the questions of whether an increased number of sample geometry instances and their abidance over many iterations of the algorithm improves the waypoint generation is of special interest to the authors.
 
@@
 
+\bibitem{hota2009}
+% Dubins Path, no wind
+% viele referenzen auf airplanes mit wind uns dubins
+Sikha Hota and Debasish Ghose, A Modified Dubins Method for Optimal Path Planning of a Miniature Air Vehicle Converging to a Straight Line Path, American Control Conference, St. Louis, MO, USA, 2009
+
+\bibitem{mcneely2007}
+% dubins vehicle und wind, wie kommeich am schnellsten von a nach b.
+Rachelle L.
+McNeely and Ram V.
+Iyer and Phillip R.
+Chandler, Tour Planning for an Unmanned Air Vehicle under Wind Conditions, Journal of Guidance, Control, and Dynamics, Vol.
+30, No.
+5, September-October 2007
+
+\bibitem{joho2007}
+Dominik Joho and Cyrill Stachniss and Patrick Pfaff and Wolfram Burgard, Autonomous Exploration for 3{D} Map Learning, {\it {A}utonome {M}obile {S}ysteme {(AMS)}}, Springer, 2007, pp. 22-28
+
 \bibitem{legrand2007}
@@
 
-\bibitem{gonzalez-banos2002}
+\bibitem{gonzalez_banos2002}
 % SLAM does not address the sensor-placement
@@
 
-\bibitem{strand}
-% Next-best-view centered
-% Grid, 2d only
-Marcus Strand and Rüdiger Dillmann, Grid based next best view planning for an autonomous robot in indoor environments
+\bibitem{strand2009}
+% Next-best-view centered. Occupancy grids are most common. cells are not binary, but carry probability.
+% grid map mit ground, wall, unexplored
+% ein gridcell ist dann guter next best view, wenn von ihm aus ein simulierter scan über möglichst viele groundcells zu unexplored gebiet führt.
+% komische metrik, was spielts für ne rolle, wieivel ground der strahl passiert hat.
+% das wäre in 3d zu teuer.
+Marcus Strand and Ruediger Dillmann, Grid based next best view planning for an autonomous robot in indoor environments
 
@@
 
+\bibitem{surmann2003}
+% Indoor only, use ICP to fuse and localize robot.
+% Next best View assumes free sensor orientation. Not possible with driving robots, but possible with mikrokopter.
+% Next best View is Art Gallery problem: Where to place guards so that they see everything.
+%  J. O?Rourke, Art Gallery Theorems and Algorithms, OxfordUniversity, Oxford, 1987
+% Man scannt, baut dann eien Grundriss und verbindet bekannte linien mit unbekannten linien.
+Die unbekannte Linie, die die meisten Überschenidungen mit laserscannerlinien hat, wird als nächstes gescannt -> highest information gain.
+% Suppose a 2D map of an art gallery is given by a polygon P, having n corners (vertices) and n walls (lines).
+If a watchman sees everything that lies on a straight line from his position, then the maximal number of guards needed is n/3 [27].
+Finding the minimum number of watchmen needed for this task is NP hard, since it can be reduced to the 3-SAT problem [27].
+Gonzalez-Banos et al.
+reduce the art gallery problem to set cover and approximate the latter one.
+Set cover is de?ned as follows: given a pair (X, F), where X is some ?nite set.
+F is a subset of the power set of X, i.e., (F ?
+P(X)) and X = s?F s.
+Find the C ?
+F , such that X = s?C C and the set C has a minimum number of elements.
+Hartmut Surmann and Andreas Nuechter and Joachim Hertzberg, An autonomous mobile robot with a 3D laser range finder for 3D exploration and digitalization of indoor environments, Robotics and Autonomous Systems 45 (2003) pp.
+181-198
+
+\bibitem{walthelm2011}
+% 2d occupancy map (float)
+% Man blickt in regionen, für die noch keine oder nur unsichere belegungshypothesen bekannt sind.
+% und dann fährt man wohl an einen punkt, von dem möglichst viele dieser regionen sichtbar sind.
+Axel Walthelm and Amir Madany Mamlouk, Multisensoric Active Spatial Environment Exploration and Modeling (2011)
+
+\bibitem{yamauchi1998}
+% 2d only
+% The central question in exploration is: Given what you know about the world, where should you move to gain as much new information as possible?
+% es gibt ein 2d grid, in dem die occupancy probability gespeichert wird.
+% dann einfach edge detection und schon hat man die frontiers.
+Brian Yamauchi, Frontier-Based Exploration Using Multiple Robots, Proceedings of the Second International Conference on Autonomous Agents, Minneanapolis, ACM Press, 1998
+
+\bibitem{shade2011}
+% occupancy grid and exploration boundaries, map besteht aus voxeln.
+Robbie Shade and Paul Newman, Choosing Where To Go: Complete 3D Exploration With Stereo, IEEE International Conference on Robotics and Automation, 2011
+
+\bibitem{orourke1987}
+Joseph O'Rourke, Art Gallery Theorems and Algorithms, Oxford University Press, Oxford, 1987
+
+\bibitem{shen2011}
+% Fliegen mit Asctec durch Innenräume, IMU und LIDAR geben Pose
+% Gutes SLAM, aber nur 2D auf mehreren Floors
+Shaojie Shen and Nathan Michael and Vijay Kumar, Autonomous Multi-Floor Indoor Navigation with a Computationally Constrained MAV, IEEE International Conference on Robotics and Automation, 2011
+
+\bibitem{triebel2006}
+Rudolph Triebel and Patrick Pfaff and Wolfram Burgard, Multi-Level Surface Maps for Outdoor Terrain Mapping and Loop Closing, Proceedings of the 2006 IEEE/RSJ International Conference on Intelligent Robots and Systems 2006
+
+\bibitem{andert2011}
+% Obstacle avoidance, path re-planning mit benzin-hubschrauber am DLR
+% Cool, aber obstacle avoidance ist noch nciht mein thema?!
+Franz Andert and Florian Adolf and Lukas Goormann and Joerg Dittrich, Mapping and Path Planning in Complex Environments: An Obstacle Avoidance Approach for an Unmanned Helicopter, IEEE International Conference on Robotics and Automation, 2011
+
 \bibitem{strand2008}
-Marcus Strand and Rüdiger Dillmann, Using an attributed 2D-grid for next-best-view planning on 3D environment data for an autonomous robot
+Marcus Strand and Ruediger Dillmann, Using an attributed 2D-grid for next-best-view planning on 3D environment data for an autonomous robot
 
-\bibitem{c2}
-H. Kwakernaak and R. Sivan, {\it Modern Signals and Systems}, Prentice Hall, Englewood Cliffs, NJ; 1991.
-
-\bibitem{c3}
-D. Boley and R. Maier, "A Parallel QR Algorithm for the Non-Symmetric Eigenvalue Algorithm", {\it in Third SIAM Conference on Applied Linear Algebra}, Madison, WI, 1988, pp. A20.
+\bibitem{xu2011}
+% Generieren sich optimale pfade, scanline-ähnlich mit holonomem uav, fixed wing.
+Anqi Xu and Chatavut Viriyasuthee and Ioannis Rekleitis, Optimal Complete Terrain Coverage Using an Unmanned Aerial Vehicle, IEEE International Conference on Robotics and Automation, 2011
 

### 2011-09-19 paper/On Physics-Based 3D Waypoint Generation/paper.tex
finale version des papers for review
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1109 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@
 \title{\LARGE \bf
-On Physics-Based 3D Waypoint Generation for Autonomous Exploration in Mobile Robotics
+On Physics-Based 3D Waypoint Generation for Autonomous Exploration in Aerial Robotics
 }
@@
         Hamburg, Germany\\
-        {\tt\small {adler,zhang}@informatik.uni-hamburg.de}%
+        {\tt\small adler@informatik.uni-hamburg.de}%
 }
@@
 This paper presents a novel, physics-based approach to computing waypoints, which are subsequently being used to steer a flying platform in three-dimensional space to map an outdoor environment.
-Because the algorithm relies on simple physical processes, it is both easy to understand and to implement in combination with traditional data structures.
-Generation of efficient sensor-trajectories for maximized information gain operates directly on unorganized point-clouds, creating a perfect fit for environment mapping with commonly user LIDAR sensors or time-of-flight cameras.
-
+Because the algorithm relies on simple physical processes, it is both easy to understand and implement in combination with traditional data structures.
+Generation of efficient sensor trajectories for maximized information gain operates directly on unorganized point clouds, creating a perfect fit for environment mapping with commonly user LIDAR sensors or time-of-flight cameras.
 We also present the algorithm by simulating its performance in a virtual outdoor scenario.
@@
 Exploration and mapping - as a subclass of this problem - has been the topic of many papers in recent years.
-The SLAM problem was first solved for two-dimensional mapping scenarios and later-on expanded to three-dimensional environments \cite{shen2011}.
+The SLAM problem was first solved for two-dimensional mapping scenarios and later-on expanded to three-dimensional environments.% \cite{shen2011}.
 
@@
 The authors of this paper, however, are implementing exploration and mapping algorithms on a flying platform with a maximum flight time of less than 15 minutes.
-Aiming to explore as much of the environment as possible within this limited time motivates an efficient way of finding sensor-poses (and in extension, -trajectories) that enable sensors to deliver as much information as quickly as possible.
-While repeatedly passing over the unknown environment in scan-line fashion results in very fast scanning, a better approach needs to be found to ensure scanning of surfaces that are shadowed from those sensor-trajectories.
+Aiming to explore as much of the environment as possible within this limited time motivates an efficient way of finding sensor poses (and in extension, sensor trajectories) that enable sensors to deliver as much information as quickly as possible.
+While repeatedly passing over the unknown environment in scan-line fashion results in very fast scanning, a better approach needs to be found to ensure scanning of surfaces that are shadowed from those sensor trajectories.
 
-This paper is organized as follows: The next section describes , investigates,
+This paper is organized as follows: the next section investigates other authors work of recent years in this and related fields.
+In the subsequent chapters, we first introduce the hardware platform used for testing our implementations and then advance to explaining the basic datastructures our algorithm relies on.
+After our algorithm is explained and illustrated, we discuss its computational complexity and the results achieved so far.
+In the final chapter, we share our thoughts on its current limitations and future developments.
 
@@
 
-Since the introduction of SLAM, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments \cite{haehnel2004}, TODO.
+Since the introduction of SLAM, much work has been done to refine and extend the algorithm for use in both two- and threedimensional environments.
 Finding the next best view, though, has never been a part of the SLAM problem \cite{gonzalez_banos2002} and is thus rarely addressed by these works.
 
-Generating safely reachable waypoints from previously acquired sensor data is an extension of next-best-view problem (which itself is an extension to the art-gallery problem \cite{orourke1987}), as it does not include the constraints of safe navigation and merely deals with the generation of isolated viewpoints.
-As sensor-poses in areas between known and unknown environment offer a good compromise between safe reachability and high information gain, frontier-based approaches are a popular method to compute new waypoints in two-dimensional space and are well documented in the literature \cite{gonzalez_banos2002}, \cite{strand2009}, \cite{yamauchi1998}, \cite{shade2011}.
+Generating safely reachable waypoints from previously acquired sensor data is an extension of next-best-view problem (which itself is an extension to the art-gallery problem \cite{orourke1987}), as it does not include the constraints of safe navigation and instead deals with the generation of isolated viewpoints\cite{chen2008}.
+As sensor poses in areas between known and unknown environment offer a good compromise between safe reachability and high information gain, frontier-based approaches are a popular method to compute new waypoints in two-dimensional space and are well documented in the literature \cite{gonzalez_banos2002}, \cite{strand2009}, \cite{yamauchi1998}, \cite{shade2011}.
 Unfortunately, information about exploration boundaries is hard to generate from three-dimensional pointclouds.
 
-As our robot is holonomic, the generated waypoints and trajectories can be created with less concern of the platform's ability to catenate them due to motion constraints that other tyes of vehicles (e.g.
+As our robot is very maneuverable, the generated waypoints and trajectories can be created with less concern of the platform's ability to catenate them due to motion constraints that other tyes of vehicles (e.g.
 fixed-wing UAVs) previously entailed \cite{hota2009}.
-The common problem of wind, appearing as both opportunity and hindrance, but mostly in the form of further constraints on feasible paths for many UAVs has spawned many contributions in the past \cite{mcneely2007}.
+The common problem of wind, manifesting as both opportunity and hindrance, but mostly in the form of further constraints on feasible paths for many UAVs has spawned many contributions in the past \cite{mcgee2005}, \cite{mcneely2007}.
 Even though wind is still a matter of concern for our platform, it does no longer influence the design of path planning or waypoint generation, but only localization and motion control.
@@
   \includegraphics[width=0.5\textwidth]{images/oktokopter_photo_small}
-  \caption{The experimental flying platform with mounted GNSS-antenna/receiver, processor-board, laserscanner and IMU.}
+  \caption{The experimental flying platform with mounted GNSS-antenna/receiver, processor-board, laser scanner and IMU.}
   \label{oktorendering}
@@
 
-To save time during the construction of our platform, we used an ``Okto 2''-octocopter from the mikrokopter-project \cite{mikrokopterproject} as a base for our vehicle (see Fig.
+%\cite{mikrokopterproject}
+To save time during the construction of our platform, we used an ``Okto 2''-octocopter from the mikrokopter-project as a base for our vehicle (see Fig.
 \ref{oktorendering}).
-Its central FlightControl (FC) processor-board is connected to eight brushless-motor controllers (BL) via an I2C bus.
+Its central FlightControl (FC) processor-board is connected to eight brushless-motor controllers (BL) via an $I{^2}C$ bus.
 Employing its on-board gyros and accelerometers, the FC is programmed to stabilize the platform by itself when no other motion-control-commands are received from either the connected remote-control-receiver or the serial port.
-The project also developed the NaviControl (NC), an add-on hardware module that sends motion-control-commands (thrust, yaw, pitch/roll each as 8-bit signed integers) to steer the platform to waypoints.
+The project also developed the NaviControl (NC), an add-on hardware module that sends motion-control-commands (thrust, yaw, pitch and roll each as 8-bit signed integers) to steer the platform to waypoints.
 We abandoned the idea of using the NC for our purposes, as its hardware adds extra weight and its control-algorithms are neither capable of moving the platform as desired nor open-source.
@@
 
-The laserscanner rotates at 2400rpm and provides both falling and rising edges after each complete scan of its 270$^\circ$ field of view.
+The laser scanner (Hokuyo UTM 30 LX) rotates at 2400rpm and provides both falling and rising edges after each complete scan of its 270$^\circ$ field of view.
 This signal is connected to the GNSS receiver, which is configured to emit the GPS Time Of Week (TOW) of the event with microsecond-precision on its USB-connection to the Atom processor-board.
-With both poses and laserscanner-events being timestamped by the GNSS-receiver's time system, we can interpolate the vehicle's pose to the scanner's pose at the time of each scan.
+With both poses and laser scanner events being timestamped by the GNSS-receiver's time system, we can interpolate the vehicle's pose to the scanner's pose at the time of each scan.
 
@@
 
-We initially planned to use two laserscanners, one with its scan-plane perpendicular to both the platform's groundplane and its forward-vector for scanning the environment and another scanner oriented to have its scan-plane parallel to the vehicle's ground-plane for collision-avoidance (see Fig.
+We initially planned to use two laser scanners, one with its scan-plane perpendicular to both the platform's groundplane and its forward-vector for scanning the environment and another scanner oriented to have its scan-plane parallel to the vehicle's ground-plane for collision-avoidance (see Fig.
 \ref{lidar_setup_pair}).
@@
 
-To avoid these drawbacks, we instead mounted a single laserscanner as depicted in Fig.
+To avoid these drawbacks, we instead mounted a single laser scanner as depicted in Fig.
 \ref{lidar_setup_single}.
-By permanently yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the world is static and the platform does not move further than the laserscanners range within one rotation.
+By permanently yawing the platform in-flight, a single scanner captures both geometry for surface reconstruction and collision avoidance, provided that the world is static and the platform does not move further than the laser scanners' range within one rotation.
 Given a constant angular yawing-speed, the disadvantage of precise interpolation of the platform's pose becomes less problematic.
@@
 
-In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based \cite{meagher1982} data-structure.
+%\cite{meagher1982}
+In order to capture sensor data representing arbitrarily complex geometry and detail, we decided to implement our algorithm on an octree-based data-structure.
 Octrees are well suited for storing information in non-uniform resolution and the process of storing and retrieving data can easily be executed in parallel, allowing for optimization of parts of our algorithms.
 Besides its position, each point in the octree also stores its brightness (i.e.
-intensity as measured by the laserscanner) and a vector pointing from that point back to the sensor.
+intensity as measured by the laser scanner) and a vector pointing from that point back to the sensor.
 As the platform's measured orientation is less precise that its measured position, this vector's length will be used lateron to assign points recorded from further distance (and thus suffering a greater impact from platform-orientation errors) a lesser weight in the following surface reconstruction.
@@
 
-\begin{figure*}[ht]
+\begin{figure*}[htb]
   \centering
-    \includegraphics[width=0.4\textwidth]{images/process_side/1}\hfill
-    \includegraphics[width=0.4\textwidth]{images/process_side/2}\vspace{3mm}
-    \includegraphics[width=0.4\textwidth]{images/process_side/3}\hfill
-    \includegraphics[width=0.4\textwidth]{images/process_side/4}
-    \caption{Waypoint generation, orthogonal projection; the bounding volume is blue, detection volume is red, vehicle position is black, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red.
-Second figure shows sampling geometry interacting with pointcloud.
-Geometry hitting the red volume will be converted to a waypoint at the position it last hit a point from the pointcloud.
-The last figure shows generated waypoints in red for an initial pointcloud population.}
+  \subfigure[]{
+    \label{method_1}
+    \includegraphics[width=0.43\textwidth]{images/process_side/1}
+  }\hfill
+  \subfigure[]{
+    \label{method_2}
+    \includegraphics[width=0.43\textwidth]{images/process_side/2}
+  }
+  \subfigure[]{
+    \label{method_3}
+    \includegraphics[width=0.43\textwidth]{images/process_side/3}
+  }\hfill
+  \subfigure[]{
+    \label{method_4}
+    \includegraphics[width=0.43\textwidth]{images/process_side/4}
+  }
+  \caption{Waypoint generation, orthogonal projection; a) depicts the initial setup: the bounding volume is blue, detection volume is red, vehicle position is black, pointcloud for geometry reconstruction is grey and pointcloud for waypoint generation is red.
+b) shows sampling geometry interacting with pointcloud.
+In c), geometry hitting the red volume below is converted to a waypoint.
+The last figure d) shows generated waypoints in red for an initial pointcloud population.
+Please see http://tinyurl.com/physwpt for an animation of the process.}
   \label{explanation_of_process}
@@
 
-Three-dimensional environment mapping is often implemented using laserscanners and time-of-flight cameras, so unorganized point clouds are a very common type of sensor data.
+Three-dimensional environment mapping is often implemented using laser scanners and time-of-flight cameras, so unorganized point clouds are a very common type of sensor data.
 Any sensor generating spatial occupancy information can be used with our algorithm, as the pointcloud is its only sensory input.
@@
 After the octree is populated with an initial set of points (i.e.
-right after lift-off), the waypoint-generation algorithm is executed as depicted in Fig.
+right after lift-off), the waypoint generation algorithm is executed as depicted in Fig.
 \ref{explanation_of_process}:
@@
 As this data structure is implemented as a map, later events caused by the same sampling geometry instance replace its previously saved position.
-     \item When a sample geometry instance collides with the detection volume, it is removed from the simulation and the position of its last collision on the ground-plane is recorded as a waypoint-marker.
-     \item Waypoint-markers are converted to waypoints if there are no other waypoints in close vicinity.
-When converting markers to waypoints, their height needs to be adjusted, as waypoint-markers tend to spawn on the scenery's ground.
+     \item When a sample geometry instance collides with the detection volume, it is removed from the simulation and the position of its last collision on the ground-plane is recorded as a waypoint marker.
+     \item Waypoint markers are converted to waypoints if there are no other waypoints in close vicinity.
+When converting markers to waypoints, their height needs to be adjusted, as waypoint markers tend to spawn on the scenery's ground.
 The distance $d$, by which markers are lifted before becoming waypoints, is calculated by $d = Range_{lidar} * \frac{SizeSampleGeom_{current}}{SizeSampleGeom_{max}}$.
-By making the distance from laserscanner to the scanned surface depend on the size of the surface's holes, a natural behavior TODO emerges: The vehicle will first scan larger voids in the surface from further away, then, as the sampling geometry's size is reduced in later iterations of the waypoint generation, come closer to fill smaller defects that couldn't be seen from former sensor poses.
+By making the distance from laser scanner to the scanned surface depend on the size of the surface's holes, a natural behavior emerges: The vehicle will first scan larger voids in the surface from further away, then, as the sampling geometry's size is reduced in later iterations of the waypoint generation, come closer to fill smaller defects that couldn't be seen from former sensor poses.
 \end{enumerate}
@@
 
+% \cite{sun2002}, 
+Depending on size, number and shape of sampling geometry, this method simulates exposing the scanned surface to a fluid pushing through the model.
+The idea of generating watertight models from previously generated datasets is rather common and already well researched \cite{hornung2006}, \cite{sormann2007}, so the novelties here are:
+
+\begin{itemize}
+ \item testing the currently collected data for watertightness in-flight and using the result as a termination criterion for exploration.
+ \item testing for watertightness using a physical simulation and its real-time implementation directly on sensor-data without additional pre-processing.
+ \item compromise between exploration time and quality of resulting model can be changed quickly and in-flight by changing sampling geometry size.
+ \item the algorithm's gracefully degraded behavior when size and number of sampling geometry cannot be scaled to achieve simulation with near-fluid behavior due to constrained computational resources.
+\end{itemize}
+
 Although any geometry can be used for waypoint generation, using spheres yields three important advantages:
+
 \begin{itemize}
@@
 
-%Spatially limited by defined exploration space, other input pre-existing sensor information and vehicle position.
-
-%Simulated annealing.
-
-\begin{figure*}[ht]
+\begin{figure*}[htb]
   \centering
-    \includegraphics[width=0.3\textwidth]{images/process_top/1}\hfill
-    \includegraphics[width=0.3\textwidth]{images/process_top/2}\hfill
-    \includegraphics[width=0.3\textwidth]{images/process_top/3}\vspace{5mm}
-    \includegraphics[width=0.3\textwidth]{images/process_top/4}\hfill
-    \includegraphics[width=0.3\textwidth]{images/process_top/5}\hfill
+  \subfigure[]{
+    \label{process_1}
+    \includegraphics[width=0.3\textwidth]{images/process_top/1}
+  }
+  \subfigure[]{
+    \label{process_2}
+    \includegraphics[width=0.3\textwidth]{images/process_top/2}
+  }
+  \subfigure[]{
+    \label{process_3}
+    \includegraphics[width=0.3\textwidth]{images/process_top/3}
+  }
+  \subfigure[]{
+    \label{process_4}
+    \includegraphics[width=0.3\textwidth]{images/process_top/4}
+  }
+  \subfigure[]{
+    \label{process_5}
+    \includegraphics[width=0.3\textwidth]{images/process_top/5}
+  }
+  \subfigure[]{
+    \label{process_6}
     \includegraphics[width=0.3\textwidth]{images/process_top/6}
+  }
   \caption{As seen from the top, multiple iterations of waypoint generation lead to continuous scanning at the border between known and unknown environment.
 Red waypoints are enqueued for scanning, green waypoints have been passed.
-In the second-to-last figure, sampling geometry of reduced size has allowed finding smaller holes in previously scanned surfaces.}
+In figures a) to d), sample spheres with $r=3m$ have been used to quickly explore the landscape, while in figure e), sampling geometry of reduced size ($r=1.5m$) has allowed finding smaller holes in previously scanned surfaces.}
   \label{figure_process}
@@
 All tests were executed in a simulator where the platform's angular velocity remained constant at $120 ^\circ/s$, while its linear velocity was limited to $2.8 m/s$ during all trials.
-After the scandata was streamed to the basestation using a TCP-connection, it was processed and statistics about flight-time and pointcloud-sizes were logged.
+After the scandata was streamed to the basestation using a TCP-connection, it was processed and statistics about flight-time and pointcloud sizes were logged.
 
@@
 
-Additionally, the graph's red line depicts our algorithms' behavior: because the pointcloud needs an initial population for waypoint generation, the scanning rate stagnates shortly after launch during the first phase of waypoint generation, then reaches comparable speeds moments later.
-As
+Additionally, the graph's red line depicts our algorithms' behavior: because the pointcloud needs an initial population for waypoint generation, the scanning rate stagnates shortly after launch during the first phase of waypoint generation, then reaches comparable speeds as the vehicle starts passing waypoints on the border between known and unknown environment.
+In this phase of flight, our algorithm is as efficient as a manually planned collision free optimal path, with the obvious advantage of having been generated automatically.
+Compared to exploring in scanline-fashion, the true strength in our method lies in the fact that on-line re-planning occurs such that when no more waypoints can be generated, gradually smaller sampling geometry is used to improve on ever smaller deficiencies in the reconstructed surface.
+This yields a truly dynamic and efficient scanning procedure that adapts well to many different kinds of outdoor scenarios.
 
@@
 With the algorithm relying on collision detection between large numbers of static and dynamic objects, it is currently limited by the speed of collision detection.
-Physics-simulation in general and collision detection especially lends itself well to optimization using massively-parallel implementations on GPUs.
+Physics-simulation in general and collision detection especially lend themselves well to optimization using massively-parallel implementations on GPUs.
 In the next steps, tests and extensions of the algorithm on GPUs are planned.
-Here, the questions of whether an increased number of sample geometry instances and their abidance over many iterations of the algorithm improves the waypoint generation is of special interest to the authors.
+Here, the questions of whether an increased number of sample geometry instances (up to a method that completely fills the user-defined scanning region with sample geometry) and their abidance over many iterations of the algorithm improves the waypoint generation is of special interest to the authors.
 
@@
 
-%\section{ACKNOWLEDGMENTS}
-%The authors gratefully acknowledge the contribution of TAMS.
+% TODO shortened
+%Finally, we plan to move our research's focus from waypoint generation to the problems of generating safely navigatable paths between them, dynamically replanning paths when necessary and discarding those that cannot be reached safely.
 
@@
 % Octrees
-\bibitem{meagher1982}
-Donald Meagher, Geometric modeling using octree encoding, {\it Computer Graphics and Image Processing}, 19, 2, pp. 129-147, (1982)
+%\bibitem{meagher1982}
+%Donald Meagher, Geometric modeling using octree encoding, {\it Computer Graphics and Image Processing}, 19, 2, pp. 129-147, (1982)
 
 \bibitem{hota2009}
+% 1mal
 % Dubins Path, no wind
@@
 \bibitem{mcneely2007}
+% 1mal
 % dubins vehicle und wind, wie kommeich am schnellsten von a nach b.
@@
 
+% 1 mal
+\bibitem{chen2008}
+Shengyong Chen and Y. F. Li and Jianwei Zhang and Wanliang Wang, Active Sensor Planning for Multiview Vision Tasks, Springer Verlag Berlin Heidelberg, 2008
+
+\bibitem{mcgee2005}
+% 1mal
+Timothy G.
+McGee and Stephen Spry and J.
+Karl Hedrick, Optimal path planning in a constant wind with a bounded turning rate, Proc.
+of the AIAA Guidance, Navigation and Control Conference and Exhibit, San Francisco, CA, August 2005.
+
 \bibitem{joho2007}
+% 2mal
 Dominik Joho and Cyrill Stachniss and Patrick Pfaff and Wolfram Burgard, Autonomous Exploration for 3{D} Map Learning, {\it {A}utonome {M}obile {S}ysteme {(AMS)}}, Springer, 2007, pp. 22-28
@@
 \bibitem{legrand2007}
+% 2 mal
 % macht broadphase per CUDA, gut erklärt in GPU gems 3, chapter 32
@@
 \bibitem{baraff1992}
+% 1mal
 David Baraff, Dynamic Simulation of Non-Penetrating Rigid Bodies, Cornell University, pp. 52 (1992)
@@
 \bibitem{gonzalez_banos2002}
+% 1mal
 % SLAM does not address the sensor-placement
@@
 
-\bibitem{haehnel2004}
+%\bibitem{haehnel2004}
 % Pioneer und Sick zu Outdoor Modellen, 16seiten, kein Wort über next-best-view
-Dirk H\"ahnel and Wolfram Burgard and Sebastian Thrun, Learning Compact 3d Models of Indoor and Outdoor Environments with a Mobile Robot, {\it Elsevier Science Special Issue Eurobot} '01, 1-16
+%Dirk H\"ahnel and Wolfram Burgard and Sebastian Thrun, Learning Compact 3d Models of Indoor and Outdoor Environments with a Mobile Robot, {\it Elsevier Science Special Issue Eurobot} '01, 1-16
 
 \bibitem{strand2009}
+% 1mal
 % Next-best-view centered. Occupancy grids are most common. cells are not binary, but carry probability.
@@
 
-\bibitem{mikrokopterproject}
-Mikrokopter project, http://www.mikrokopter.de, 2011
+%\bibitem{mikrokopterproject}
+%Mikrokopter project, http://www.mikrokopter.de, 2011
 
-\bibitem{surmann2003}
+%\bibitem{surmann2003}
 % Indoor only, use ICP to fuse and localize robot.
@@
 F , such that X = s?C C and the set C has a minimum number of elements.
-Hartmut Surmann and Andreas Nuechter and Joachim Hertzberg, An autonomous mobile robot with a 3D laser range finder for 3D exploration and digitalization of indoor environments, Robotics and Autonomous Systems 45 (2003) pp.
+%Hartmut Surmann and Andreas Nuechter and Joachim Hertzberg, An autonomous mobile robot with a 3D laser range finder for 3D exploration and digitalization of indoor environments, Robotics and Autonomous Systems 45 (2003) pp.
 181-198
@@
 \bibitem{walthelm2011}
+% 2mal
 % 2d occupancy map (float)
@@
 \bibitem{yamauchi1998}
+% 2mal
 % 2d only
@@
 \bibitem{shade2011}
+% 2mal
 % occupancy grid and exploration boundaries, map besteht aus voxeln.
@@
 \bibitem{orourke1987}
+% 1mal
 Joseph O'Rourke, Art Gallery Theorems and Algorithms, Oxford University Press, Oxford, 1987
 
-\bibitem{shen2011}
+%\bibitem{shen2011}
 % Fliegen mit Asctec durch Innenräume, IMU und LIDAR geben Pose
 % Gutes SLAM, aber nur 2D auf mehreren Floors
-Shaojie Shen and Nathan Michael and Vijay Kumar, Autonomous Multi-Floor Indoor Navigation with a Computationally Constrained MAV, IEEE International Conference on Robotics and Automation, 2011
+%Shaojie Shen and Nathan Michael and Vijay Kumar, Autonomous Multi-Floor Indoor Navigation with a Computationally Constrained MAV, IEEE International Conference on Robotics and Automation, 2011
 
 \bibitem{triebel2006}
-Rudolph Triebel and Patrick Pfaff and Wolfram Burgard, Multi-Level Surface Maps for Outdoor Terrain Mapping and Loop Closing, Proceedings of the 2006 IEEE/RSJ International Conference on Intelligent Robots and Systems 2006
+% 1mal
+%Rudolph Triebel and Patrick Pfaff and Wolfram Burgard, Multi-Level Surface Maps for Outdoor Terrain Mapping and Loop Closing, Proceedings of the 2006 IEEE/RSJ International Conference on Intelligent Robots and Systems 2006
+Rudolph Triebel et al., Multi-Level Surface Maps for Outdoor Terrain Mapping and Loop Closing, Proceedings of the 2006 IEEE/RSJ IROS 2006
 
-\bibitem{andert2011}
+%\bibitem{andert2011}
+% 0mal
 % Obstacle avoidance, path re-planning mit benzin-hubschrauber am DLR
 % Cool, aber obstacle avoidance ist noch nciht mein thema?!
-Franz Andert and Florian Adolf and Lukas Goormann and Joerg Dittrich, Mapping and Path Planning in Complex Environments: An Obstacle Avoidance Approach for an Unmanned Helicopter, IEEE International Conference on Robotics and Automation, 2011
+%Franz Andert and Florian Adolf and Lukas Goormann and Joerg Dittrich, Mapping and Path Planning in Complex Environments: An Obstacle Avoidance Approach for an Unmanned Helicopter, IEEE International Conference on Robotics and Automation, 2011
 
-\bibitem{strand2008}
-Marcus Strand and Ruediger Dillmann, Using an attributed 2D-grid for next-best-view planning on 3D environment data for an autonomous robot
+%\bibitem{strand2008}
+% 0mal
+%Marcus Strand and Ruediger Dillmann, Using an attributed 2D-grid for next-best-view planning on 3D environment data for an autonomous robot
 
 \bibitem{xu2011}
+% 1mal
 % Generieren sich optimale pfade, scanline-ähnlich mit holonomem uav, fixed wing.
 Anqi Xu and Chatavut Viriyasuthee and Ioannis Rekleitis, Optimal Complete Terrain Coverage Using an Unmanned Aerial Vehicle, IEEE International Conference on Robotics and Automation, 2011
+
+%\bibitem{sun2002}
+% 1mal, unwichtig
+% machen 3d reconstruction und sind stolz auf watertightness.
+%Y. Sun and J. K. Paik and A. Koschan and M. A. Abidi, 3D Reconstruction of Indoor and Outdoor Scenes Using a Mobile Range Scanner
+
+\bibitem{sormann2007}
+% 1mal
+% watertight reconstruction
+Mario Sormann and Christopher Zach and Joachim Bauer and Konrad Karner and Horst Bishof, Watertight Multi-view Reconstruction Based on Volumetric Graph-Cuts
+
+\bibitem{hornung2006}
+% 1mal
+% watertight reconstruction
+Alexander Hornung and Leif Kobbelt, Robust Reconstruction of Watertight 3D Models from Non-uniformly Sampled Point Clouds Without Normal Information
 

### 2011-09-19 basestation/basestation.cpp
stand nach paper einreichung
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1111 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@ -30,6 +30,26 @@ BaseStation::BaseStation() : QMainWindow()
     connect(mTcpSocket, SIGNAL(connected()), SLOT(slotSocketConnected()));
     connect(mTcpSocket, SIGNAL(disconnected()), SLOT(slotSocketDisconnected()));
 
+
+
+    mTimerStats = new QTimer(this);
+    mTimerStats->setInterval(1000);
+    mTimerStats->start();
+
+    //mDateTimeProgramStart = QDateTime::currentDateTime();
+
+    mStatsFile = new QFile(QDateTime::currentDateTime().toString("yyyyMMdd-hhmmsszzz").prepend("stats-").append(".txt"));
+    if(!mStatsFile->open(QIODevice::WriteOnly | QIODevice::Text))
+    {
+        QMessageBox::critical(this, "File Error", QString("Couldn't open stats-file %1 for writing!").arg(mStatsFile->fileName()));
+    }
+    QTextStream out(mStatsFile);
+    out << "# FlightTime;ItemsInOctreeFine;ItemsInOctreeCoarse\n";
+    out.flush();
+    connect(mTimerStats, SIGNAL(timeout()), SLOT(slotWriteStats()));
+
+
+
     mControlWidget = new ControlWidget(this);
     connect(this, SIGNAL(vehiclePoseChanged(Pose)), mControlWidget, SLOT(slotUpdatePose(Pose)));
     addDockWidget(Qt::RightDockWidgetArea, mControlWidget);
@@ -55,18 +75,22 @@ BaseStation::BaseStation() : QMainWindow()
     connect(mControlWidget, SIGNAL(wayPointSwap(quint16,quint16)), mFlightPlanner, SLOT(slotWayPointSwap(quint16,quint16)));
 
     connect(mFlightPlanner, SIGNAL(wayPointDeleted(quint16)), mControlWidget, SLOT(slotWayPointDeleted(quint16)));
-    connect(mFlightPlanner, SIGNAL(wayPointInserted(quint16,WayPoint)), mControlWidget, SLOT(slotWayPointInserted(quint16,WayPoint)));
     connect(mFlightPlanner, SIGNAL(wayPoints(QList<WayPoint>)), mControlWidget, SLOT(slotSetWayPoints(QList<WayPoint>)));
[…]

### 2011-09-19 basestation/basestation.h

@@ -56,6 +56,10 @@ private:
 
     GlWidget *mGlWidget;
 
+    QTimer *mTimerStats;
+    QFile *mStatsFile;
+    QDateTime mDateTimeLastLidarInput;
+
     void addRandomPoint();
     void keyPressEvent(QKeyEvent* event);
 
@@ -80,12 +84,15 @@ private slots:
     void slotConnectToRover(void);
     void slotExportCloud(void);
 
-    void slotWayPointInserted(const quint16&, const WayPoint&);
-    void slotWayPointDeleted(const quint16&);
-    void slotSetWayPoints(const QList<WayPoint>&);
+    void slotRoverWayPointInsert(const quint16&, const WayPoint&);
+    void slotRoverWayPointDelete(const quint16&);
+    void slotRoverWayPointsSet(const QList<WayPoint>&);
 
     void slotSendRtkDataToRover(const QByteArray& rtkData);
     void slotSendData(const QByteArray &data);
+
+    void slotWriteStats();
+    void slotAddLogFileMarkForPaper(QList<WayPoint> wptList);
 
 public:
     BaseStation(void);

### 2011-09-19 basestation/controlwidget.cpp

@@ -18,6 +18,8 @@ ControlWidget::ControlWidget(BaseStation* baseStation) : QDockWidget((QWidget*)b
     connect(mBtnWptPrepend, SIGNAL(clicked()), SLOT(slotWayPointPrepend()));
     connect(mBtnWptAppend, SIGNAL(clicked()), SLOT(slotWayPointAppend()));
     connect(mBtnWptDelete, SIGNAL(clicked()), SLOT(slotWayPointDelete()));
+    connect(mBtnWptLoad, SIGNAL(clicked()), SLOT(slotWayPointLoad()));
+    connect(mBtnWptSave, SIGNAL(clicked()), SLOT(slotWayPointSave()));
 
     connect(mBtnWptUp, SIGNAL(clicked()), SLOT(slotWayPointUp()));
     connect(mBtnWptDown, SIGNAL(clicked()), SLOT(slotWayPointDown()));
@@ -328,4 +330,81 @@ void ControlWidget::slotWayPointsCleared()
 {
     mWayPointTable->clear();
     mWayPointTable->setRowCount(0);
+}
+
+void ControlWidget::slotWayPointLoad()
+{
+    const QString fileName = QFileDialog::getOpenFileName(this, "Load WayPoints");
+    if(!fileName.isNull())
+    {
+        QFile file(fileName);
+        if(!file.open(QIODevice::ReadOnly | QIODevice::Text))
+        {
+            QMessageBox::critical(this, "File Error", "Couldn't open file for reading!");
+            return;
+        }
+
+        int i = 0;
+        while (!file.atEnd())
+        {
+            const QString line(file.readLine());
+            const QStringList values = line.split(";", QString::SkipEmptyParts);
+
+            if(values.size() != 3)
+            {
+                QMessageBox::warning(this, "File Format Error", QString("Not three numbers in line %1!").arg(i));
+                return;
+            }
+
+            const WayPoint wpt(
+                        QVector3D(
+                            values.at(0).toFloat(),
+                            values.at(1).toFloat(),
+                            values.at(2).toFloat()
+                            )
+                        );
+
+            emit wayPointInsert(mWayPointTable->rowCount(), wpt);
+
+            i++;
+        }
+
+        file.close();
+    }
+}
+
+void ControlWidget::slotWayPointSave()
+{
+    const QString fileName = QFileDialog::getSaveFileName(this, "Save WayPoints");
+    if(!fileName.isNull())
+    {
[…]

### 2011-09-19 basestation/controlwidget.h

@@ -38,6 +38,8 @@ public slots:
     void slotUpdateRtkStatus(bool working = false);
     void slotUpdateRoverConnection(bool connected);
 
+    void slotSetWayPointCoordinateFields(Qt::MouseButton, QVector3D);
+
     // Called by FlightPlanner when it has changed its internal list.
     void slotWayPointInserted(const quint16& index, const WayPoint& waypoint);
     void slotWayPointDeleted(const quint16& index);
@@ -48,13 +50,16 @@ private slots:
     void slotSimulationStarted();
     void slotSimulationPaused();
 
-    // Called when buttons are pressed, simply emit signals to be processed my FlightPlanner
+    // Called when buttons are pressed, simply emit signals to be processed by FlightPlanner
     void slotWayPointPrepend();
     void slotWayPointAppend();
     void slotWayPointDelete();
     void slotWayPointChange(int, int);
     void slotWayPointUp();
     void slotWayPointDown();
+
+    void slotWayPointLoad();
+    void slotWayPointSave();
 
     void slotSetScanVolume();
 

### 2011-09-19 basestation/flightplannerinterface.cpp

@@ -22,16 +22,16 @@ void FlightPlannerInterface::slotSetScanVolume(const QVector3D min, const QVecto
 
 void FlightPlannerInterface::sortToShortestPath(QList<WayPoint> &wayPoints, const QVector3D &currentVehiclePosition)
 {
-    qDebug() << "FlightPlannerInterface::sortToShortestPath(): vehicle is at" << currentVehiclePosition;
+//    qDebug() << "FlightPlannerInterface::sortToShortestPath(): vehicle is at" << currentVehiclePosition;
 
     float distanceBefore = 0;
     for(int i=1;i<wayPoints.size();i++) distanceBefore += wayPoints.at(i-1).distanceToLine(wayPoints.at(i), QVector3D());
-    qDebug() << "FlightPlannerInterface::sortToShortestPath(): total distance between" << wayPoints.size() << "points before:" << distanceBefore;
+//    qDebug() << "FlightPlannerInterface::sortToShortestPath(): total distance between" << wayPoints.size() << "points before:" << distanceBefore;
 
     QList<WayPoint> wps(wayPoints);
     float distanceBeforewps = 0;
     for(int i=1;i<wps.size();i++) distanceBeforewps += wps.at(i-1).distanceToLine(wps.at(i), QVector3D());
-    qDebug() << "FlightPlannerInterface::sortToShortestPath(): wps total distance between" << wps.size() << "points before:" << distanceBeforewps;
+//    qDebug() << "FlightPlannerInterface::sortToShortestPath(): wps total distance between" << wps.size() << "points before:" << distanceBeforewps;
 
     wayPoints.clear();
     wayPoints.append(currentVehiclePosition);
@@ -59,7 +59,7 @@ void FlightPlannerInterface::sortToShortestPath(QList<WayPoint> &wayPoints, cons
 
     float distanceAfter = 0;
     for(int i=1;i<wayPoints.size();i++) distanceAfter += wayPoints.at(i-1).distanceToLine(wayPoints.at(i), QVector3D());
-    qDebug() << "FlightPlannerInterface::sortToShortestPath(): total distance between" << wayPoints.size() << "points after:" << distanceAfter;
+//    qDebug() << "FlightPlannerInterface::sortToShortestPath(): total distance between" << wayPoints.size() << "points after:" << distanceAfter;
 }
 
[…]

### 2011-09-19 basestation/flightplannerinterface.h

@@ -19,7 +19,7 @@ protected:
     QWidget* mParentWidget;
     QList<WayPoint>* mWayPointsAhead, *mWayPointsPassed;
 
-    static void sortToShortestPath(QList<WayPoint> &wayPoints, const QVector3D &currentVehiclePosition);
+    static void sortToShortestPath(QList<WayPoint> &wayPointsSetOnRover, const QVector3D &currentVehiclePosition);
 
 public:
     FlightPlannerInterface(QWidget* widget, Octree* pointCloud);
@@ -39,6 +39,7 @@ public:
 public slots:
     void slotWayPointDelete(const quint16& index);
     void slotWayPointInsert(const quint16& index, const WayPoint& wpt);
+    void slotWayPointInsertedByRover(const quint16& index, const WayPoint& wpt);
     void slotWayPointSwap(const quint16& i, const quint16& j);
     void slotWayPointsClear();
 
@@ -46,13 +47,23 @@ public slots:
 
     virtual void slotSetScanVolume(const QVector3D min, const QVector3D max);
     virtual void slotGenerateWaypoints() = 0;
-    virtual void slotVisualize() const = 0;
+    virtual void slotVisualize() const;
 
     void slotVehiclePoseChanged(const Pose& pose);
 
 signals:
+    // Emitted to tell other classes that waypoint @wpt was inserted at index @index
     void wayPointInserted(const quint16& index, const WayPoint& wpt);
+
+    // Emitted to tell the rover that it should insert waypoint @wpt at index @index
+    void wayPointInsertOnRover(const quint16& index, const WayPoint& wpt);
+
+    // Emitted to tell other classes that waypoint @index was deleted. We do not care how or by whom
     void wayPointDeleted(const quint16& index);
+    // Emitted to tell the rover that it should delete waypoint @index
+    void wayPointDeleteOnRover(const quint16& index);
+
+    void wayPointsSetOnRover(QList<WayPoint>);
     void wayPoints(QList<WayPoint>);
 
     void suggestVisualization();

### 2011-09-19 basestation/flightplannerphysics.cpp

@@ -15,7 +15,8 @@ FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, Octree* pointCloud) 
     connect(mDialog, SIGNAL(deleteSampleGeometry()), SLOT(slotDeleteSampleGeometry()));
     connect(mDialog, SIGNAL(processPhysics(bool)), SLOT(slotProcessPhysics(bool)));
     connect(mDialog, SIGNAL(createSafePath()), SLOT(slotCreateSafePathToNextWayPoint()));
-    connect(mDialog, SIGNAL(submitWayPoints()), SLOT(slotEmitWayPoints()));
+    connect(mDialog, SIGNAL(submitWayPoints()), SLOT(slotSubmitGeneratedWayPoints()));
+    connect(mDialog, SIGNAL(deleteWayPoints()), SLOT(slotDeleteGeneratedWayPoints()));
 
     connect(mDialog, SIGNAL(gravityChanged(QVector3D)), SLOT(slotGravityChanged(QVector3D)));
     connect(mDialog, SIGNAL(frictionChanged(float,float)), SLOT(slotFrictionChanged(float,float)));
@@ -51,9 +52,9 @@ FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, Octree* pointCloud) 
 
     mDeletionTriggerShape = new btBoxShape(
                 btVector3( // These are HALF-extents!
-                    (mScanVolumeMax.x() - mScanVolumeMin.x()) / 2.0,
+                    (mScanVolumeMax.x() - mScanVolumeMin.x()) / 2,
                     10,
-                    (mScanVolumeMax.z() - mScanVolumeMin.z()) / 2.0));
+                    (mScanVolumeMax.z() - mScanVolumeMin.z()) / 2));
 
     mGhostObjectDeletionTrigger = new btGhostObject;
     mGhostObjectDeletionTrigger->setWorldTransform(mTransformDeletionTrigger);
@@ -87,7 +88,7 @@ FlightPlannerPhysics::FlightPlannerPhysics(QWidget* widget, Octree* pointCloud) 
     mGhostObjectVehicle->setRestitution(0.5);
     mGhostObjectVehicle->setWorldTransform(mTransformVehicle);
     mGhostObjectVehicle->setCollisionShape(mBodyVehicle->getCollisionShape());
[…]

### 2011-09-19 basestation/flightplannerphysics.h

@@ -23,6 +23,8 @@ public:
 
     void insertPoint(LidarPoint* const point);
 
+    qint64 getNumberOfPointsInCollisionOctree();
+
 private:
     Octree* mOctree;
     FlightPlannerPhysicsDialog* mDialog;
@@ -85,7 +87,8 @@ private slots:
     void slotCreateSampleGeometry();
     void slotDeleteSampleGeometry();
     void slotProcessPhysics(bool);
-    void slotEmitWayPoints();
+    void slotSubmitGeneratedWayPoints();
+    void slotDeleteGeneratedWayPoints();
     void slotGravityChanged(const QVector3D& gravity);
     void slotFrictionChanged(const float& frictionPointCloudGround, const float& frictionSampleGeometry);
     void slotRestitutionChanged(const float& restitutionPointCloudGround, const float& restitutionSampleGeometry);

### 2011-09-19 basestation/flightplannerphysicsdialog.cpp

@@ -13,6 +13,7 @@ FlightPlannerPhysicsDialog::FlightPlannerPhysicsDialog(QWidget *parent) :
     connect(ui->mBtnCreateSafePath, SIGNAL(clicked()), SIGNAL(createSafePath()));
 
     connect(ui->mBtnSubmitWaypoints, SIGNAL(clicked()), SIGNAL(submitWayPoints()));
+    connect(ui->mBtnDeleteWaypoints, SIGNAL(clicked()), SIGNAL(deleteWayPoints()));
 
     connect(ui->mSpinBoxGravityX, SIGNAL(valueChanged(double)), SLOT(slotGravityChanged()));
     connect(ui->mSpinBoxGravityY, SIGNAL(valueChanged(double)), SLOT(slotGravityChanged()));

### 2011-09-19 basestation/flightplannerphysicsdialog.h

@@ -51,6 +51,7 @@ signals:
     void frictionChanged(float,float);
     void restitutionChanged(float,float);
     void submitWayPoints();
+    void deleteWayPoints();
 };
 
 #endif // FLIGHTPLANNERPHYSICSDIALOG_H

### 2011-09-19 basestation/glwidget.cpp

@@ -75,12 +75,17 @@ void GlWidget::initializeGL()
 //    glEnable(GL_DEPTH_TEST);
 //    glDepthFunc(GL_LEQUAL);
     */
-    glHint(GL_PERSPECTIVE_CORRECTION_HINT, GL_NICEST);
+//    glHint(GL_PERSPECTIVE_CORRECTION_HINT, GL_NICEST);
     /* we use resizeGL once to set up our initial perspective */
 //    resizeGL(width, height);
     /* Reset the rotation angle of our object */
 //    rotQuad = 0;
 //    glFlush();
+
+    // needed to convert mous epos to world pos, see
+    // http://stackoverflow.com/questions/3089271/converting-mouse-position-to-world-position-opengl
+    glEnable(GL_DEPTH);
+
 
 
     glShadeModel(GL_SMOOTH);						// Enable Smooth Shading
@@ -108,8 +113,10 @@ void GlWidget::resizeGL(int w, int h)
     glViewport(0, 0, w, h);
     glMatrixMode(GL_PROJECTION);
     glLoadIdentity();
-    gluPerspective(50.0*mZoomFactor, (GLfloat)w/(GLfloat)h, 10, +8000.0);
+//    gluPerspective(50.0*mZoomFactor, (GLfloat)w/(GLfloat)h, 10, +8000.0);
 //    gluOrtho2D(-w/2.0, w/2.0, -h/2.0, h/2.0);
+//    gluOrtho2D(-200.0, 200.0, -200.0, 200.0);
+    glOrtho(-150 * mZoomFactor, 150 * mZoomFactor, -150 * mZoomFactor, 150 * mZoomFactor, 1, 10000);
     glTranslatef(camPos.x(), camPos.y(), camPos.z());
     glMatrixMode(GL_MODELVIEW);
 }
@@ -126,7 +133,6 @@ void GlWidget::moveCamera(const QVector3D &pos)
     glMatrixMode(GL_MODELVIEW);
     */
 }
-
 
 void GlWidget::paintGL()
 {
@@ -191,6 +197,24 @@ void GlWidget::paintGL()
     glTranslatef(-mCamLookAt.x(), -mCamLookAt.y(), -mCamLookAt.z());
 
 
+
+    // Draw base plate for unProjecting
+    /*glEnable(GL_BLEND);
+        glDisable(GL_LIGHTING);
[…]

### 2011-09-19 basestation/glwidget.h

@@ -48,6 +48,7 @@ class GlWidget : public QGLWidget
     void zoom(double zoomFactor);
     void drawAxes(const GLfloat& x, const GLfloat& y, const GLfloat& z, const GLfloat& red, const GLfloat& green, const GLfloat& blue) const;
     void drawVehicleVelocity() const;
+    QVector3D convertMouseToWorldPosition(const QPoint&);
 
 public:
     GlWidget(BaseStation *baseStation, Octree* octree, FlightPlannerInterface* flightPlanner);
@@ -64,6 +65,7 @@ protected:
 
 signals:
     void visualizeNow();
+    void mouseClickedAtWorldPos(Qt::MouseButton, QVector3D);
 
 public slots:
     void slotViewFromTop();

### 2011-09-19 common/baseconnection.cpp

@@ -99,15 +99,17 @@ void BaseConnection::processPacket(QByteArray packet)
         WayPoint wayPoint;
         stream >> index;
         stream >> wayPoint;
-        emit wayPointInsert(index, wayPoint);
         qDebug() << "BaseConnection::processPacket(): inserting waypoint from base to index" << index;
+        emit wayPointInsert(index, wayPoint);
+        qDebug() << "BaseConnection::processPacket(): inserting waypoint from base to index" << index << "Done.";
     }
     else if(command == "waypointdelete")
     {
         quint16 index;
         stream >> index;
+        qDebug() << "BaseConnection::processPacket(): base tells me to delete wpt" << index << "will now emit wptDelete";
         emit wayPointDelete(index);
-        qDebug() << "BaseConnection::processPacket(): deleting waypoint" << index << "from base.";
+        qDebug() << "BaseConnection::processPacket(): base tells me to delete wpt" << index << "will now emit wptDelete. Done.";
     }
     else if(command == "waypoints")
     {
@@ -115,6 +117,7 @@ void BaseConnection::processPacket(QByteArray packet)
         stream >> wayPointList;
         qDebug() << "BaseConnection::processPacket(): received" << wayPointList.size() << "waypoints from base.";
         emit wayPoints(wayPointList);
+        qDebug() << "BaseConnection::processPacket(): received" << wayPointList.size() << "waypoints from base. Done.";
     }
 //    else if(command == "getstatus")
 //    {
@@ -227,16 +230,18 @@ void BaseConnection::slotFlushWriteQueue()
 
 // called when the rover has changed the waypoints list, will be sent to base
 // Disabled: Why should the rover ever send waypoints?
-//void BaseConnection::slotNewWayPointsFromRover(const QVector<WayPoint>& wayPoints)
-//{
-//    qDebug() << "sending new waypoints to base:" << wayPoints;
-//    QByteArray data;
-//    QDataStream stream(&data, QIODevice::WriteOnly);
+// Update: Because it can cretae its own ladning-waypoint, the base will want to know!
+void BaseConnection::slotRoverWayPointInserted(const quint16& index, const WayPoint& wayPoint)
+{
+    qDebug() << "sending new waypoints to base:" << wayPoint;
+    QByteArray data;
+    QDataStream stream(&data, QIODevice::WriteOnly);
 
-//    stream << QString("waypoints");
-//    stream << wayPoints;
[…]

### 2011-09-19 common/baseconnection.h

@@ -72,8 +72,8 @@ signals:
 //    void statusRequested();
 
 public slots:
-    // called when the rover has changed the waypoints list, will be sent to base
-//    void slotNewWayPointsFromRover(const QVector<WayPoint>& wayPoints);
+    // called when the rover has appended a waypoint, will be sent to base
+    void slotRoverWayPointInserted(const quint16& index, const WayPoint& wayPoint);
 
     // called by rover when it has reached a waypoint, notifies basestation
     void slotWayPointReached(const WayPoint&wpt);

### 2011-09-19 common/flightcontroller.cpp

@@ -127,10 +127,12 @@ void FlightController::slotComputeMotionCommands()
         double errorHeight = nextWayPoint.y() - mLastKnownVehiclePose.position.y();
         mErrorIntegralHeight += errorHeight*timeDiff;
         double derivativeHeight = (errorHeight - mPrevErrorHeight + 0.00001)/timeDiff;
-        double outputThrust = (Kp*errorHeight) + (0.5*Ki*mErrorIntegralHeight) + (Kd*derivativeHeight);
+        double outputThrust = (Kp*errorHeight) + (3.0*0.5*Ki*mErrorIntegralHeight) + (Kd*derivativeHeight);
 
         outputPitch /= 10.0;
         outputRoll /= 10.0;
+
+        qDebug() << "no wpts" << mWayPoints.size() << "next wpt height" << nextWayPoint.y() << "curr height" << mLastKnownVehiclePose.position.y() << "thrust" << outputThrust;
 
 //        qDebug() << "values PRYH:" << QString::number(mLastKnownVehiclePose.getPitchDegrees(), 'f', 2) << "\t" << QString::number(mLastKnownVehiclePose.getRollDegrees(), 'f', 2) << "\t" << QString::number(mLastKnownVehiclePose.getYawDegrees(), 'f', 2) << "\t" << QString::number(mLastKnownVehiclePose.position.y(), 'f', 2);
 //        qDebug() << "should PRYH:" << QString::number(desiredPitch, 'f', 2) << "\t" << QString::number(desiredRoll, 'f', 2) << "\t" << QString::number(mDesiredYaw, 'f', 2) << "\t" << QString::number(nextWayPoint.y(), 'f', 2);
@@ -196,25 +198,28 @@ void FlightController::wayPointReached()
     // The current waypoint has been reached.
     mWayPointsPassed.append(mWayPoints.takeFirst());
 
+    qDebug() << "FlightController::wayPointReached(): reached waypoint" << mWayPointsPassed.last();
+
     Q_ASSERT(getFlightState() == ApproachingNextWayPoint);
 
     if(mWayPoints.size())
     {
         emit message(QString("%1::%2(): ").arg(metaObject()->className()).arg(__FUNCTION__), Information, "waypoint reached, more waypoints present, approaching");
+        qDebug() << "FlightController::wayPointReached(): reached waypoint, approaching next";
     }
     else if(mLastKnownBottomBeamLength < 0.2)
     {
         setFlightState(Idle);
[…]

### 2011-09-19 simulator/physics.cpp

@@ -43,6 +43,8 @@ Physics::Physics(Simulator *simulator, OgreWidget *ogreWidget) :
     else
         mVehicleNode = mOgreWidget->createVehicleNode("vehicleNode", Ogre::Vector3(10, 100, 10), Ogre::Quaternion::IDENTITY);
 
+    mVehicleNode->setPosition(140,90,150);
+
     mVehicleNode->attachObject(mVehicleEntity);
 
     mEngineNodes.append(mVehicleNode->createChildSceneNode(Ogre::Vector3(+0.00, +0.00, -0.20), Ogre::Quaternion(Ogre::Degree(000), Ogre::Vector3(1, 0, 0))));  // engine 1, forward, CW
@@ -458,10 +460,10 @@ void Physics::slotUpdateWind()
         btVector3 windVector = mVectorWind.at(sampleNumber);
         mVehicleBody->applyCentralForce(windVector * mWindFactor);
 
-        qDebug() << "Physics::slotUpdateWind(): applying wind sample" << sampleNumber << ":" << windVector.x() << windVector.y() << windVector.z() << "and factor" << mWindFactor;
+//        qDebug() << "Physics::slotUpdateWind(): applying wind sample" << sampleNumber << ":" << windVector.x() << windVector.y() << windVector.z() << "and factor" << mWindFactor;
     }
-    else
-        qDebug() << "Physics::slotUpdateWind(): not applying wind: enabled:" << mWindEnable << "factor" << mWindFactor << "samples:" << mVectorWind.size();
+//    else
+//        qDebug() << "Physics::slotUpdateWind(): not applying wind: enabled:" << mWindEnable << "factor" << mWindFactor << "samples:" << mVectorWind.size();
 }
 
 float Physics::getHeightAboveGround()

### 2011-09-19 simulator/ogrewidget.cpp
fix texturing, no more blakc buildings
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1112 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@ -848,7 +848,7 @@ void OgreWidget::setupTerrain()
 
     entity = mSceneManager->createEntity("tudorHouse", "tudorhouse.mesh");
     entity->setQueryFlags(0xFFFFFFFF);
-    rotation.FromAngleAxis(Ogre::Degree(Ogre::Math::RangeRandom(-180, 180)), Ogre::Vector3::UNIT_Y);
+    rotation.FromAngleAxis(Ogre::Degree(Ogre::Math::RangeRandom(-180, 180)-20), Ogre::Vector3::UNIT_Y);
     sceneNode = mSceneManager->getRootSceneNode()->createChildSceneNode(
                 position + Ogre::Vector3(0, mTerrainGroup->getHeightAtWorldPosition(position) + mTerrainPos.y + 6.5, -13),
                 rotation);
@@ -858,7 +858,7 @@ void OgreWidget::setupTerrain()
 
     entity = mSceneManager->createEntity("church", "church.mesh");
     entity->setQueryFlags(0xFFFFFFFF);
-    rotation.FromAngleAxis(Ogre::Degree(Ogre::Math::RangeRandom(-180, 180)), Ogre::Vector3::UNIT_Y);
+    rotation.FromAngleAxis(Ogre::Degree(Ogre::Math::RangeRandom(-180, 180)+180.0), Ogre::Vector3::UNIT_Y);
     sceneNode = mSceneManager->getRootSceneNode()->createChildSceneNode(
                 position + Ogre::Vector3(-5, mTerrainGroup->getHeightAtWorldPosition(position) + mTerrainPos.y + 0.0, 55),
                 rotation);
@@ -877,7 +877,7 @@ void OgreWidget::setupTerrain()
     entity = mSceneManager->createEntity("house2", "house2.mesh");
     entity->setMaterialName("house2/8_-_Default");
     entity->setQueryFlags(0xFFFFFFFF);
-    rotation.FromAngleAxis(Ogre::Degree(Ogre::Math::RangeRandom(-180, 180)), Ogre::Vector3::UNIT_Y);
+    rotation.FromAngleAxis(Ogre::Degree(Ogre::Math::RangeRandom(-180, 180)-100), Ogre::Vector3::UNIT_Y);
     sceneNode = mSceneManager->getRootSceneNode()->createChildSceneNode(
[…]

### 2011-09-21 common/pose.cpp
jaja
git-svn-id: https://politbuero.dyndns.org/svn/repos/promotion@1113 f4953187-ecf7-0310-90f3-8d0fedb0aae5

@@ -52,8 +52,10 @@ Pose Pose::interpolateLinear(const Pose &before, const Pose &after, const float 
 }
 
 // http://paulbourke.net/miscellaneous/interpolation/
+
+// TODO: This code doesn't really use the timestamps in the poses, that seems stupid
 Pose Pose::interpolateCubic(const Pose * const first, const Pose * const before, const Pose * const after, const Pose * const last, const float &mu)
-{ //                                      y0                 y1                 y2                 y3
+{//                                             y0                        y1                         y2                        y3
     Q_ASSERT(mu <= 0.0 && mu <= 1.0);
 
     const double mu2 = mu*mu;
@@ -169,7 +171,8 @@ Pose Pose::operator+(const Pose &p) const
                 keepWithinRangeRadians(mYaw + p.getYawRadians()),
                 keepWithinRangeRadians(mPitch + p.getPitchRadians()),
                 keepWithinRangeRadians(mRoll + p.getRollRadians()),
-                (timestamp + p.timestamp)/2
+                // use the latest timestamp, needed in LaserScanner::slotNewVehiclePose(const Pose& pose)
+                std::max(timestamp,p.timestamp)
                 );
 }
 

### 2011-09-21 koptercontrol/gpsdevice.cpp

@@ -539,6 +539,13 @@ void GpsDevice::processSbfData()
 
                 emit newVehiclePose(p);
 
+                // Write log data: timestamp[space]V1[space]V2[space]...[space]Vn\n
+                QTextStream out(&mLogFile);
+                out << timestamp;
+                std::vector<long>::iterator itr;
+                for(itr=mScanDistancesNext.begin();itr != mScanDistancesNext.end(); ++itr) out << " " << *itr;
+                out << "\n";
+
                 mPoseClockDivisor = mPoseClockDivisor % 20;
                 if(mPoseClockDivisor % 20 == 0) emit newVehiclePoseLowFreq(p);
 
@@ -562,7 +569,7 @@ void GpsDevice::processSbfData()
             qDebug() << "SBF: ExtEvent";
             if(msgIdRev != 2)
             {
-                qWarning() << "GpsDevice::processSbfData(): WARNING: invalid revision" << msgIdRev << "for block id" << msgIdBlock;
+                qWarning() << "GpsDevice::processSbfData(): WARNING: invalid revision" << msgIdRev << "for block id" << msgIdBlock << "will not emit scanFinished!";
                 break;
             }
             // process

### 2011-09-21 koptercontrol/koptercontrol.cpp

@@ -84,15 +84,13 @@ KopterControl::KopterControl(int argc, char **argv) : QCoreApplication(argc, arg
     connect(mKopter, SIGNAL(kopterStatus(quint32, qint16, float)), mBaseConnection, SLOT(slotNewVehicleStatus(quint32, qint16, float)));
 
     connect(mLaserScanner, SIGNAL(bottomBeamLength(const float&)), mFlightController, SLOT(slotSetBottomBeamLength(const float&)));
-
-    //connect(mRtkFetcher, SIGNAL(rtkData(const QByteArray&)), mGpsDevice, SLOT(slotSetRtkData(const QByteArray&)));
-
     connect(mBaseConnection, SIGNAL(enableScanning(const bool&)), mLaserScanner, SLOT(slotEnableScanning(const bool&)));
     connect(mBaseConnection, SIGNAL(rtkDataReady(const QByteArray&)), mGpsDevice, SLOT(slotSetRtkData(const QByteArray&)));
 
     connect(mGpsDevice, SIGNAL(newVehiclePose(Pose)), mFlightController, SLOT(slotSetVehiclePose(Pose)));
     connect(mGpsDevice, SIGNAL(newVehiclePose(Pose)), mLaserScanner, SLOT(slotNewVehiclePose(Pose)));
     connect(mGpsDevice, SIGNAL(newVehiclePoseLowFreq(Pose)), mBaseConnection, SLOT(slotPoseChanged(Pose)));
+    connect(mGpsDevice, SIGNAL(scanFinished(quint32)), mLaserScanner, SLOT(slotScanFinished(quint32)));
 
     connect(mFlightController, SIGNAL(motion(quint8,qint8,qint8,qint8,qint8)), mKopter, SLOT(slotSetMotion(quint8,qint8,qint8,qint8,qint8)));
 

### 2011-09-21 koptercontrol/laserscanner.cpp

@@ -7,6 +7,10 @@ LaserScanner::LaserScanner(const QString &deviceFileName, const Pose &pose)
     mDeviceFileName = deviceFileName;
     mRelativePose = pose;
 
+    mLogFile = new QFile(QDateTime::currentDateTime().toString("yyyyMMdd-hhmmsszzz").prepend("scannerdata-").append(".log"));
+    if(!mLogFile->open(QIODevice::WriteOnly | QIODevice::Text))
+        qFatal("LaserScanner::LaserScanner(): Couldn't open logfile %s for writing, exiting.", qPrintable(mLogFile->fileName()));
+
     mIsEnabled = false;
 
     mPoints.reserve(1200); // should be 1080 points (270 degrees, 0.25° resolution)
@@ -68,7 +72,7 @@ float LaserScanner::getHeightAboveGround() const
     const int index = mScanner.deg2index(90);
     if(mScanDistancesCurrent->size() > index)
     {
-        return (float)((*mScanDistancesCurrent)[index]) - 0.22;
+        return (float)((*mScanDistancesCurrent)[index]) - 0.21;
     }
     else
     {
@@ -93,6 +97,13 @@ void LaserScanner::slotScanFinished(const quint32 &timestamp)
         long timestamp = 0;
         if(mScanner.capture(*mScanDistancesNext, &timestamp) <= 0)
             qWarning() << "LaserScanner::slotScanFinished(): weird, less than 1 samples sent by lidar";
+
+        // Write log data: scan[space]timestamp[space]V1[space]V2[space]...[space]Vn\n
+        QTextStream out(&mLogFile);
+        out << "scan " << timestamp;
+        std::vector<long>::iterator itr;
+        for(itr=mScanDistancesNext.begin();itr != mScanDistancesNext.end(); ++itr) out << " " << *itr;
+        out << "\n";
 
         if(mScannerPoseFirst != 0 && mIsEnabled)
         {
@@ -140,8 +151,9 @@ void LaserScanner::slotNewVehiclePose(const Pose& pose)
     QMutexLocker locker(&mMutex);
     qDebug() << "LaserScanner::slotNewVehiclePose(): received a pose from gps-device at time" << pose.timestamp;
 
-    // The hokuyo finished a scan and sent a falling edge to the GpsDevice, which now notifies us of
-    // this scan together with the pose in which scanning finished.
[…]

### 2011-09-21 koptercontrol/laserscanner.h

@@ -21,6 +21,8 @@ class LaserScanner : public QObject//QThread
 private:
     QString mDeviceFileName;
 
+    QFile* mLogFile;
+
     bool mIsEnabled;
 
     UrgCtrl mScanner;

### 2011-09-21 simulator/ogrewidget.cpp

@@ -158,6 +158,15 @@ void OgreWidget::keyPressEvent(QKeyEvent *e)
         update();
         e->accept();
     }
+    else if(e->key() == Qt::Key_M)
+    {
+        if(mCamera->getPolygonMode() == Ogre::PM_SOLID)
+            mCamera->setPolygonMode(Ogre::PM_WIREFRAME);
+        else
+            mCamera->setPolygonMode(Ogre::PM_SOLID);
+
+        update();
+    }
     else
     {
         e->ignore();
@@ -933,7 +942,7 @@ void OgreWidget::configureTerrainDefaults(Ogre::Light* light)
     // CompositeMapDistance decides how far the Ogre terrain will render the lightmapped terrain.
     mTerrainGlobals->setCompositeMapDistance(3000);
 
-    //mTerrainGlobals->setUseRayBoxDistanceCalculation(true);
+//    mTerrainGlobals->setUseRayBoxDistanceCalculation(true);
     //mTerrainGlobals->getDefaultMaterialGenerator()->setDebugLevel(1);
     //mTerrainGlobals->setLightMapSize(256);
 

### 2011-09-21 simulator/physics.cpp

@@ -412,7 +412,7 @@ void Physics::slotSetMotion(const quint8& thrust, const qint8& pitch, const qint
 
 bool Physics::initializeWind()
 {
-    QFile windDataFile("winddata-10Hz.txt");
+    QFile windDataFile("winddata.txt");
     if(!windDataFile.open(QIODevice::ReadOnly | QIODevice::Text))
     {
         emit windInitialized(false);
